---
title: "Elastic cylinder modal series solution"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Elastic cylinder modal series solution}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  status = c("experimental", "unvalidated"),
  pages = c(
    Overview = "index.html",
    Implementation = "ecms-implementation.html",
    Theory = "ecms-theory.html"
  )
)
```

The elastic-cylinder modal-series solution is available through `target_strength(..., model = "ecms")`. The preferred geometry carrier is an elastic-cylinder `ESS` object, while the elastic material parameters are supplied through:

- `density_body`
- `sound_speed_longitudinal_body`
- `sound_speed_transversal_body`

No separate public local ECMS package was identified during this pass, so the implementation check here uses an independent direct transcription of the Faran-Stanton algebra rather than a package-to-package comparison.

# Reference case

The reference cylinder uses:

- length `40 mm`
- radius `5 mm`
- body density `2800 kg m^-3`
- longitudinal speed `6398 m s^-1`
- transverse speed `3122 m s^-1`
- surrounding water density `1026.8 kg m^-3`
- surrounding water sound speed `1477.3 m s^-1`
- broadside incidence
- `12-200 kHz` in `2 kHz` steps

In `acousticTS`, the call is:

```{r}
library(acousticTS)

elastic_cylinder <- fls_generate(
  shape = cylinder(
    length_body = 0.04,
    radius_body = 0.005,
    n_segments = 201
  ),
  density_body = 2800,
  sound_speed_body = 1500,
  theta_body = pi / 2
)

elastic_cylinder <- target_strength(
  elastic_cylinder,
  frequency = seq(12e3, 200e3, by = 2e3),
  model = "ecms",
  density_sw = 1026.8,
  sound_speed_sw = 1477.3,
  sound_speed_longitudinal_body = 6398,
  sound_speed_transversal_body = 3122
)

head(extract(elastic_cylinder, "model")$ECMS)
```

# Summary

```{r echo = FALSE}
ecms_summary <- utils::read.csv(
  file.path("..", "..", "scratch", "ecms_reference_compare_summary.csv")
)

knitr::kable(
  ecms_summary,
  digits = 6,
  col.names = c(
    "Comparison",
    "Max abs. delta TS (dB)",
    "Mean abs. delta TS (dB)",
    "f at max delta (kHz)",
    "acousticTS elapsed (s)",
    "Reference elapsed (s)"
  )
)
```

The direct transcription and the package output coincide exactly on the shared grid.

## Spectrum overlay

```{r fig.width = 8, fig.height = 6}
ecms_compare <- utils::read.csv(
  file.path("..", "..", "scratch", "ecms_reference_compare.csv")
)
ecms_compare <- ecms_compare[order(ecms_compare$frequency_hz), ]

par(mfrow = c(2, 1), mar = c(4, 4.5, 2, 1))

plot(
  ecms_compare$frequency_khz,
  ecms_compare$TS_reference,
  type = "l",
  lwd = 2.5,
  col = "black",
  xlab = "",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2))
)
lines(ecms_compare$frequency_khz, ecms_compare$TS_acousticts, col = "#0b6e4f", lwd = 2)
legend(
  "bottomleft",
  legend = c("Direct reference", "acousticTS"),
  col = c("black", "#0b6e4f"),
  lwd = c(2.5, 2),
  bty = "n"
)

plot(
  ecms_compare$frequency_khz,
  ecms_compare$TS_acousticts - ecms_compare$TS_reference,
  type = "l",
  lwd = 2,
  col = "#0b6e4f",
  xlab = "Frequency (kHz)",
  ylab = expression(Delta * TS ~ (dB))
)
abline(h = 0, lty = 3, col = "grey50")
```

# Closing note

This comparison is intentionally literal. The goal here is not to compare one software package to another. It is to show that the code path in `acousticTS` reproduces the elastic-cylinder algebra it claims to implement, across a full frequency band rather than only at a few checkpoint frequencies.
