---
title: "Implementation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{SPHMS implementation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

The `acousticTS` package uses object-based scatterers so the same implementation pattern carries across models: create a scatterer, run `target_strength()`, inspect the stored model output, and then compare a small set of physically important inputs. For SPHMS, the object must be spherical and the main model choice is the boundary condition.

That apparent simplicity is one of the reasons SPHMS is so useful. When the target really is well approximated by a sphere, the geometry and the coordinate system match exactly, so the main practical decisions collapse to size, material contrast, and boundary interpretation rather than to complicated numerical bookkeeping.

## Sphere object generation

The example below uses a gas-filled sphere so the same geometry can be reused for gas-filled and idealized rigid or pressure-release comparisons.

```{r}
library(acousticTS)

sphere_shape <- sphere(
  radius_body = 5e-3,
  n_segments = 80
)

gas_sphere <- gas_generate(
  shape = sphere_shape,
  g_fluid = 0.0012,
  h_fluid = 0.22,
  theta_body = pi / 2
)

gas_sphere
```

This setup is a good example of the geometry-versus-physics distinction that runs through the package. The outer geometry is simply spherical. The scatterer constructor then gives that sphere a specific physical interpretation, here a gas-filled interior with strong density and sound-speed contrast relative to the surrounding fluid.

## Calculating a TS-frequency spectrum

```{r}
frequency <- seq(18e3, 120e3, by = 6e3)

gas_sphere <- target_strength(
  object = gas_sphere,
  frequency = frequency,
  model = "sphms",
  boundary = "gas_filled"
)
```

This call is mechanically simple, but it is still worth reading carefully. The `model` argument selects the spherical modal-series solution, while the `boundary` argument chooses which spherical boundary-value problem is being solved. In other words, SPHMS does not mean only one thing acoustically. The same spherical geometry can be used under different interface assumptions, and that choice is often the most important physical decision in the run.

## Extracting model results

Model results can be extracted either visually or directly through `extract()`.

### Plotting results

```{r fig.width = 6}
plot(gas_sphere, type = "shape")
plot(gas_sphere, type = "model")
```

### Accessing results

```{r}
sphms_results <- extract(gas_sphere, "model")$SPHMS
head(sphms_results)
```

At this stage, readers should confirm that the returned output matches the intended spherical problem. A gas-filled run should usually look materially different from a rigid or pressure-release run of the same size, especially when the contrast is strong. This is also a good point to confirm that the frequency grid and reporting variables are the ones needed for the later comparison.

## Comparing boundary conditions

Because the geometry is fixed, boundary-condition comparisons are a clean way to understand what part of the response is due to the interface assumption rather than the size of the sphere itself.

```{r fig.width = 8, fig.height = 4.5}
rigid_sphere <- fls_generate(
  shape = sphere_shape,
  density_body = 1045,
  sound_speed_body = 1520,
  theta_body = pi / 2
)

pressure_release_sphere <- fls_generate(
  shape = sphere_shape,
  density_body = 1045,
  sound_speed_body = 1520,
  theta_body = pi / 2
)

rigid_sphere <- target_strength(
  rigid_sphere,
  frequency,
  "sphms",
  boundary = "fixed_rigid"
)

pressure_release_sphere <- target_strength(
  pressure_release_sphere,
  frequency,
  "sphms",
  boundary = "pressure_release"
)

ts_rigid <- extract(rigid_sphere, "model")$SPHMS
ts_pressure_release <- extract(pressure_release_sphere, "model")$SPHMS
ts_gas <- extract(gas_sphere, "model")$SPHMS

par(oma = c(0, 0.25, 0, 0), mar = c(5, 6, 4, 2))
plot(
  x = ts_rigid$frequency * 1e-3,
  y = ts_rigid$TS,
  type = "l",
  lty = 1,
  lwd = 2.5,
  xlab = "Frequency (kHz)",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  cex.lab = 1.3,
  cex.axis = 1.2
)
lines(x = ts_pressure_release$frequency * 1e-3, y = ts_pressure_release$TS, col = "royalblue4", lty = 1, lwd = 2.5)
lines(x = ts_gas$frequency * 1e-3, y = ts_gas$TS, col = "firebrick3", lty = 1, lwd = 2.5)
legend(
  "bottomright",
  legend = c("Fixed rigid", "Pressure release", "Gas filled"),
  col = c("black", "royalblue4", "firebrick3"),
  lwd = 2.5,
  lty = c(1, 1, 1),
  cex = 1.0,
  bty = "n"
)
```

This comparison is especially helpful because it isolates the role of the boundary condition cleanly. The outer radius is the same in all three runs. What changes is the interface physics: suppressed normal motion, vanishing surface pressure, or a highly compressible gas-filled interior. That makes the resulting differences much easier to interpret than if geometry and boundary were changed simultaneously.

For fluid-filled spheres, the same implementation pattern applies; the main change is the selected `boundary` and the interior material properties. In practical SPHMS work, the first controls to revisit are usually the sphere radius, the selected boundary condition, and the medium-to-interior contrasts, because those are the inputs that determine the modal coefficient problem most directly.
