---
title: "Model library"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Model library}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

The family library is easiest to navigate when read against model-comparison reviews, organism-focused scattering surveys, and companion open-source modeling ecosystems [@jech_etal_2015; @stanton_acoustic_1996; @brandyn_lucca_osoceanacousticsechopop_2026].



The package now contains enough model families that the site works better when the model pages are treated as a library rather than as a flat list in the top navigation. This page is the main entry point for that library.

## Model-status policy

```{r model_status_policy, echo=FALSE, results='asis'}
acousticTS:::.validation_status_policy()
```

```{r model_library_listing, echo=FALSE, results='asis'}
acousticTS:::.validation_model_library()
```

# How to use this page

Each family page is intended to be the local navigation hub for that model family. Use the family page to move between overview, theory, implementation, and related package pages, rather than relying on a long global theory or implementation menu.
