---
output: rmarkdown::github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, 
                      comment="##", 
                      fig.retina=2, 
                      fig.path = "README_figs/README-")
```

# acousticTS

Acoustic backscatter from a single target or organism is expressed as
the intensity of an echo typically denoted as the *backscattering
cross-section* (σ~bs~, m^2^). Target strength (TS, dB re. 1 m^2^) is the
logarithmic representation of σ~bs~ where: TS = 10 log~10~ (σ~bs~). TS
can be used to convert integrated (e.g. nautical area scattering
coefficient, S~A~, dB re. 1(m^2^ nmi^-2^) or volumetric backscatter
(e.g. S~v~, dB re. 1 m^-1^) collected from fisheries acoustic surveys
into units of number density, such as the volumetric density of a fish
school (i.e. animals m^-3^). This parameter can also aid in classifying
backscatter based on the multifrequency response of targets, such as
separating likely echoes of large predatory fish from smaller prey.
While there are several approaches for estimating TS, one common method
is to apply physics-based models to predict theoretical TS that comprise
exact and approximate solutions. The models provided in the `acousticTS`
package can help provide TS estimates parameterized using broad
statitsical distributions of inputs. This package is in a constant state
of development with updates to the available model library,
computational efficiency, and quality-of-life improvements.

*General DOI* <https://doi.org/10.5281/zenodo.7600659>

*Latest release DOI*
[![DOI](https://zenodo.org/badge/161965429.svg)](https://zenodo.org/badge/latestdoi/161965429)

## [Installation]{.underline}

You can install the current released version of acousticTS via:

``` r
devtools::install_github("brandynlucca/acousticTS")
```

Or you can install the development version of acousticTS like so:

``` r
devtools::install_github("brandynlucca/acousticTS@test-branch")
```

## [Models currently available]{.underline}

#### Two-ray model for uniformly bent fluid-like cylinders ([DCM]{.underline})

*Stanton, T.K., Clay, C.S., and Chu, D. (1993). Ray representation of
sound scattering by weakly scattering deformed fluid cylinders: Simple
physics and application to zooplankton. J. Acoust. Soc. Am., 94,
3454-3462.*

#### Distorted wave Born approximation ([DWBA]{.underline})

*Stanton, T.K., Chu, D., and Wiebe, P.H. (1998). Sound scattering by
several zooplankton groups. II. Scattering models. J. Acoust. Soc. Am.,
103, 236-253.*

#### Stochastic distorted wave Born approximation ([SDWBA]{.underline})

*Demer, D.A., and Conti, S.G. 2003. Reconciling theoretical versus
empirical target strengths of krill: effects of phase variability on the
distorted-wave Born approximation. ICES J. Mar. Sci., 60, 429-434.*

#### SDWBA for uniformly bent scatterers ([SDWBA_curved]{.underline})

*Demer, D.A., and Conti, S.G. 2003. Reconciling theoretical versus
empirical target strengths of krill: effects of phase variability on the
distorted-wave Born approximation. ICES J. Mar. Sci., 60, 429-434.*

*Stanton, T.K., Chu, D., and Wiebe, P.H. (1998). Sound scattering by
several zooplankton groups. II. Scattering models. J. Acoust. Soc. Am.,
103, 236-253.*

#### Kirchoff-ray mode approximation ([KRM]{.underline})

*Clay C.S. and Horne J.K. (1994). Acoustic models of fish: The Atlantic
cod (Gadus morhua). J. Acoust. Soc. Am., 96, 1661-1668.*

#### Modal series solution for gas-filled fluid spheres ([MSS_anderson]{.underline})

*Anderson, V.C. (1950). Sound scattering from a fluid sphere. J. Acoust.
Soc. Am., 22, 426-431.*

#### Homogeneous solid sphere ([calibration]{.underline})

*MacLennan D. N. (1981). The theory of solid spheres as sonar
calibration targets. Scottish Fisheries Research No. 22, Department of
Agriculture and Fisheries for Scotland.*

## [Examples]{.underline}

Below are examples of different models used to predict TS for a sardine
with a gas-filled swimbladder, a tungsten carbide calibration sphere, a
crustacean modeled as a prolate spheroid, and a generic gas-filled
bubble.

### Kirchoff Ray-Mode approximation for a Sardine with a gas-filled swimbladder

```{r sardine}
library( acousticTS )
### Call in the built-in sardine shape dataset
data( sardine )
### Inspect the object
print( sardine )
plot( sardine )
### We will now define a frequency range to predict TS over
frequency <- seq( 1e3 , 400e3 , 1e3 )
### And now we use the target_strength(...) function to model TS for this fish
sardine <- target_strength( sardine, 
                            frequency = frequency, 
                            model = "KRM" )
### Plot results
plot( sardine, type = 'model' )
### Extract model results
sardine_ts <- extract( sardine , "model") $KRM
```

### Calibration sphere

```{r calibration}
### Let's create a calibration sphere 
### Default inputs here are a 38.1 mm diameter and a tungsten carbide 
### (WC) material properties.
cal_sphere <- cal_generate( )
### We will use the same frequency range as the previous example
### Calculate TS
cal_sphere <- target_strength( object = cal_sphere,
                               frequency = frequency,
                               model = "calibration" )
### Plot results
plot( cal_sphere , type = 'model' )
### Extract model results
calibration_ts <- extract( cal_sphere , "model" )$calibration
```

### Fluid sphere (Anderson, 1950)

```{r bubble}
### Let's create a gas-filled bubble with a raidus of 4 mm
### This defaults to a density contrast, g_body, of 0.0012
### This defaults to a soundspeed contrast, h_body, of 0.220
bubble <- gas_generate( radius = 4e-3 ,
                        ID = "gas bubble" )
print( bubble )
### Model TS using the Anderson (1950) model
bubble <- target_strength( bubble,
                           frequency = seq( 1e3 , 300e3 , 0.5e3 ) ,
                           model = "MSS_anderson" )
### Plot results
plot( bubble , type = 'model' )
```

### Fluid-like crustacean (prolate spheroid) using the distorted Born wave approximation (DWBA) and ray-based deformed cylinder model (DCM)

```{r crustacean}
### First let's create a prolate spheroid shape 
### 25 mm long with a length-to-radius ratio of 16
crustacean <- fls_generate( shape = "prolate_spheroid" ,
                            length_body = 25.0e-3 , 
                            length_radius_ratio = 16 ,
                            radius_curvature_ratio = 3.3 ,
                            g_body = 1.03 ,
                            h_body = 1.02 )
print( crustacean )
plot( crustacean )
### Model TS using the ray-path deformed cylinder model (DCM), distorted wave
### Born approximation (DWBA), the stochastic variation of the DWBA (SDWBA) , 
### and specifically curved versions of both the DWBA and SDWBA 
crustacean <- target_strength( crustacean ,
                               frequency = seq( 1e3 , 200e3 , 1e3 ) ,
                               model = c( "DCM" , "DWBA" , "SDWBA" ,
                                          "DWBA_curved" , "SDWBA_curved" ) )
### Plot results
plot( crustacean , type = 'model' )
```
