---
title: "Implementation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Elastic-shelled sphere implementation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

The `acousticTS` package uses object-based scatterers so the same implementation pattern carries across models: create a scatterer, run `target_strength()`, inspect the stored model output, and then compare a small set of physically important inputs. For `ESSMS`, the required object class is `ESS`, which combines a spherical shell, an optional internal fluid, and the elastic constants required for the shell solution.

## Elastic-shelled sphere object generation

An `ESS` object can be created with `ess_generate()`. For the implementation below, the shell is described by its outer radius and thickness, while the shell material is described by density, sound speed, and elastic constants. The inner fluid can be provided using either contrasts or absolute material properties.

```{r}
library(acousticTS)

shelled_sphere <- ess_generate(
  shape = "sphere",
  radius_shell = 10e-3,
  shell_thickness = 0.8e-3,
  density_shell = 2565,
  sound_speed_shell = 3750,
  density_fluid = 1077.3,
  sound_speed_fluid = 1575,
  E = 7.0e10,
  nu = 0.32
)

shelled_sphere
```

## Calculating a TS-frequency spectrum

The `target_strength()` wrapper initializes the ESSMS model, performs the modal calculation, and stores the results back inside the same object. As with the rest of the package, frequency is supplied in Hz.

```{r}
frequency <- seq(38e3, 120e3, by = 4e3)

shelled_sphere <- target_strength(
  object = shelled_sphere,
  frequency = frequency,
  model = "essms"
)
```

## Extracting model results

Model results can be extracted either visually or directly through `extract()`.

### Plotting results

```{r fig.width = 6}
plot(shelled_sphere, type = "shape")
# plot(shelled_sphere, type = "model")
```

### Accessing results

```{r}
essms_results <- extract(shelled_sphere, "model")$ESSMS
head(essms_results)
```

The extracted `data.frame` contains the modeled frequency, complex backscattering amplitude `f_bs`, backscattering cross-section `sigma_bs`, and target strength `TS`.

## Comparing shell thicknesses

Shell thickness strongly affects the resonance structure of the ESSMS solution, so it is a natural first comparison when testing a new parameterization.

```{r fig.width = 8, fig.height = 4.5}
sphere_thin <- ess_generate(
  shape = "sphere",
  radius_shell = 10e-3,
  shell_thickness = 0.4e-3,
  density_shell = 2565,
  sound_speed_shell = 3750,
  density_fluid = 1077.3,
  sound_speed_fluid = 1575,
  E = 7.0e10,
  nu = 0.32
)

sphere_thick <- ess_generate(
  shape = "sphere",
  radius_shell = 10e-3,
  shell_thickness = 1.2e-3,
  density_shell = 2565,
  sound_speed_shell = 3750,
  density_fluid = 1077.3,
  sound_speed_fluid = 1575,
  E = 7.0e10,
  nu = 0.32
)

sphere_thin <- target_strength(sphere_thin, frequency, "essms")
sphere_thick <- target_strength(sphere_thick, frequency, "essms")

ts_base <- extract(shelled_sphere, "model")$ESSMS
ts_thin <- extract(sphere_thin, "model")$ESSMS
ts_thick <- extract(sphere_thick, "model")$ESSMS

par(oma = c(0, 0.25, 0, 0), mar = c(5, 6, 4, 2))
plot(
  x = ts_base$frequency * 1e-3,
  y = ts_base$TS,
  type = "l",
  lty = 1,
  lwd = 2.5,
  xlab = "Frequency (kHz)",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  cex.lab = 1.3,
  cex.axis = 1.2
)
lines(x = ts_thin$frequency * 1e-3, y = ts_thin$TS, col = "royalblue4", lty = 1, lwd = 2.5)
lines(x = ts_thick$frequency * 1e-3, y = ts_thick$TS, col = "firebrick3", lty = 1, lwd = 2.5)
legend(
  "bottomright",
  legend = c("0.8 mm shell", "0.4 mm shell", "1.2 mm shell"),
  col = c("black", "royalblue4", "firebrick3"),
  lwd = 2.5,
  lty = c(1, 1, 1),
  cex = 1.0,
  bty = "n"
)
```

When you move from a tutorial object to a real calibration or biological shell, the next quantities to revisit are the shell elastic constants and the shell-to-fluid property contrast, because those control where the strongest modal features occur.
