---
title: "Running Models"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Running Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

[`target_strength()`](../../reference/target_strength.html) is the common
interface for running models in `acousticTS`. It combines a scatterer, a
frequency vector, and one or more model names. The function returns the
scatterer with model parameters and results attached.

This page covers execution and inspection. Use [Choosing a
Model](../model-selection/model-selection.html) for geometry, boundary, and
validity considerations. Consult the model's theory and implementation pages
for its assumptions and numerical controls.

::: {.warning data-title="A completed calculation is not a validation result"}
Successful execution confirms that the supplied inputs passed the implemented
checks. It does not establish that the model is appropriate or validated for
the target and parameter range. Review [Validation and
Benchmarks](../validation-benchmarks/validation-benchmarks.html) before using a
model outside its documented scope.
:::

# Available models

[`available_models()`](../../reference/available_models.html) lists built-in and
user-registered models, their accepted aliases, result names, and source:

```{r eval=FALSE}
library(acousticTS)

available_models()
```

Model names are case-insensitive. An alias and its canonical name resolve to the
same registered model. Availability does not imply compatibility with every
scatterer or shape.

# Run one model

Build the scatterer first, then supply frequencies in hertz and a model name:

```{r}
library(acousticTS)

shape_obj <- cylinder(
  length_body = 0.03,
  radius_body = 0.0025,
  n_segments = 60
)

scatterer_obj <- fls_generate(
  shape = shape_obj,
  density_body = 1045,
  sound_speed_body = 1520,
  theta_body = pi / 2,
  ID = "example cylinder"
)

frequency <- seq(38e3, 120e3, by = 6e3)

scatterer_obj <- target_strength(
  object = scatterer_obj,
  frequency = frequency,
  model = "dwba"
)
```

Reassignment is required. `target_strength()` returns an updated S4 object. It
does not modify the object bound to `scatterer_obj` unless the returned value is
assigned.

The function performs two stages for each model. The initializer converts the
scatterer and call arguments into the model-specific parameterization. The
solver then calculates and stores the result. Use `show()` for a compact object
summary, then inspect the two result slots with
[`extract()`](../../reference/extract.html):

```{r}
show(scatterer_obj)
names(extract(scatterer_obj, "model_parameters"))
names(extract(scatterer_obj, "model"))

dwba_output <- extract(scatterer_obj, "model")$DWBA
names(dwba_output)
head(dwba_output)
```

Output fields vary by model. Inspect their names before assuming that every
result contains the same amplitude, cross-section, acoustic-size, or diagnostic
columns.

# Model-specific arguments

Arguments after `model` are forwarded to the selected model initializer. Their
meaning is model-specific. Examples include boundary conditions, medium
properties, approximation methods, integration settings, and stochastic
controls. Use the model reference page to determine which are required.

```{r eval=FALSE}
target_strength(
  object = sphere_obj,
  frequency = frequency,
  model = "sphms",
  boundary = "gas_filled"
)

target_strength(
  object = fls_obj,
  frequency = frequency,
  model = "hpa",
  method = "stanton"
)
```

`verbose = TRUE` prints model initialization and execution progress. It is
useful for locating the stage at which a larger call fails:

```{r eval=FALSE}
target_strength(
  object = scatterer_obj,
  frequency = frequency,
  model = "dwba",
  verbose = TRUE
)
```

# Run several models

A vector of model names runs the requested models sequentially on the same
scatterer:

```{r}
comparison_obj <- fls_generate(
  shape = shape_obj,
  density_body = 1045,
  sound_speed_body = 1520,
  theta_body = pi / 2
)

comparison_obj <- target_strength(
  object = comparison_obj,
  frequency = frequency,
  model = c("dwba", "hpa")
)

names(extract(comparison_obj, "model"))
```

This keeps geometry, material properties, and frequency fixed while different
model formulations are evaluated. It does not make the models physically
equivalent. Compare only outputs with compatible definitions and domains.

## Shared and per-model arguments

Arguments supplied through `...` are shared with the requested model
initializers when relevant. Use `model_args` for arguments that belong to one
model. A model-specific value overrides a shared value with the same name:

```{r eval=FALSE}
comparison_obj <- target_strength(
  object = scatterer_obj,
  frequency = frequency,
  model = c("dwba", "sdwba"),
  density_sw = 1026,
  sound_speed_sw = 1478,
  model_args = list(
    sdwba = list(
      n_iterations = 100,
      n_segments_init = 14,
      phase_sd_init = sqrt(2) / 2,
      length_init = 38.35e-3,
      frequency_init = 120e3
    )
  )
)
```

Each `model_args` name must identify a requested model. Each entry must be a
named list or named atomic vector.

# Plot and inspect results

The returned object retains its target definition and the stored model output.
[`plot()`](../../reference/plot.Scatterer.html) can display either:

```{r fig.width=6, fig.height=4, fig.cap="Geometry stored on the modeled scatterer."}
plot(scatterer_obj, type = "shape")
```

```{r fig.width=6, fig.height=4, fig.cap="Model output stored on the same scatterer."}
plot(scatterer_obj, type = "model")
```

For quantitative work, extract the underlying result rather than reading values
from a plot. Target strength is appropriate for logarithmic reporting.
Cross-section or complex amplitude may be required for linear averaging,
coherent combination, or mechanism-level analysis. Before comparison or export,
confirm the shape and orientation, requested model, frequency grid, and output
columns.

# When to run again

Treat each stored result as belonging to one run configuration. Run the model
again after changing:

- geometry, segmentation, or component placement
- material properties, contrasts, or orientation
- frequency or other acoustic conditions
- model name, boundary condition, method, or numerical controls

Create curved targets with [`brake()`](../../reference/brake.html) before
running the relevant model.

# Execution checks

Before interpreting a result, confirm:

- the scatterer class and shape are supported by the model
- frequencies are in hertz and use the intended grid
- medium and target properties use consistent units and definitions
- model-specific arguments represent the intended physical and numerical case
- results are stable under relevant resolution or truncation changes
- compared models use the same target, conditions, and reporting quantity

Continue with [Comparing Models](../comparing-models/comparing-models.html) or
[Simulation and Parameter Sweeps](../simulation-parameter-sweeps/simulation-parameter-sweeps.html).
