---
title: "Phase-compensated distorted wave Born approximation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Phase-compensated distorted wave Born approximation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  family = "pcdwba",
  pages = c(
    Overview = "index.html",
    Theory = "pcdwba-theory.html",
    Implementation = "pcdwba-implementation.html",
    Benchmarks = "pcdwba-benchmarks.html",
    Validation = "pcdwba-validation.html"
  )
)
```


These pages follow the phase-compensated weak-scattering literature for broadside elongated bodies and krill-style applications [@Chu_1999; @Chu_1993].

The phase-compensated distorted wave Born approximation is available through `target_strength(..., model = "pcdwba")`. The implementation is intended for weakly scattering fluid-like bodies and uses the same curved-cylinder bookkeeping whether the target starts as a canonical bent cylinder or an arbitrary fluid-like profile.

This page uses a single bent-cylinder case to show how the phase-compensated weak-scattering calculation is built, run, and inspected.

::: {.note data-title="Reference case"}
The same bent-cylinder case is used on the benchmark page and in the baseline validation row, so object setup, timing, and direct-integral residuals can be read against the same geometry.
:::

## Reference case

The comparison uses a single reproducible bent-cylinder case:

- length `15 mm`
- radius `1 mm`
- taper order `10`
- curvature ratio `rho_c / L = 3`
- density contrast `g = 1.02`
- sound-speed contrast `h = 1.02`
- broadside incidence
- `12-200 kHz` in `2 kHz` steps
- `51` integration nodes

In acousticTS, that target is built as:

```{r}
library(acousticTS)

pcdwba_object <- fls_generate(
  shape = cylinder(
    length_body = 0.015,
    radius_body = 0.001,
    taper = 10,
    radius_curvature_ratio = 3,
    n_segments = 50
  ),
  g_body = 1.02,
  h_body = 1.02,
  theta_body = pi / 2
)

pcdwba_object <- target_strength(
  object = pcdwba_object,
  frequency = seq(12e3, 200e3, by = 2e3),
  model = "pcdwba",
  sound_speed_sw = 1500,
  density_sw = 1026
)

head(extract(pcdwba_object, "model")$PCDWBA)
```

## Related result pages

For numerical result tables, use:

- [PCDWBA benchmarks](pcdwba-benchmarks.html)
- [PCDWBA validation](pcdwba-validation.html)
