---
title: "Phase-compensated distorted wave Born approximation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Phase-compensated distorted wave Born approximation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  status = c("experimental", "validated"),
  pages = c(
    Overview = "index.html",
    Implementation = "pcdwba-implementation.html",
    Theory = "pcdwba-theory.html"
  )
)
```

The phase-compensated distorted wave Born approximation is available through `target_strength(..., model = "pcdwba")`. The implementation is intended for weakly scattering fluid-like bodies and uses the same curved-cylinder bookkeeping whether the target starts as a canonical bent cylinder or an arbitrary fluid-like profile.

This page checks the implementation against two source-level references:

- the `pcdwba_fbs` routine in the local `echopop` source tree,
- the bent-cylinder DWBA routines in the local `ZooScatR` source tree.

# Reference case

The comparison uses a single reproducible bent-cylinder case:

- length `15 mm`
- radius `1 mm`
- taper order `10`
- curvature ratio `rho_c / L = 3`
- density contrast `g = 1.02`
- sound-speed contrast `h = 1.02`
- broadside incidence
- `12-200 kHz` in `2 kHz` steps
- `51` integration nodes in all three implementations

In `acousticTS`, that target is built as:

```{r}
library(acousticTS)

pcdwba_object <- fls_generate(
  shape = cylinder(
    length_body = 0.015,
    radius_body = 0.001,
    taper = 10,
    radius_curvature_ratio = 3,
    n_segments = 50
  ),
  g_body = 1.02,
  h_body = 1.02,
  theta_body = pi / 2
)

pcdwba_object <- target_strength(
  object = pcdwba_object,
  frequency = seq(12e3, 200e3, by = 2e3),
  model = "pcdwba",
  sound_speed_sw = 1500,
  density_sw = 1026
)

head(extract(pcdwba_object, "model")$PCDWBA)
```

# Comparison summary

```{r echo = FALSE}
pcdwba_summary <- utils::read.csv(
  file.path("..", "..", "scratch", "pcdwba_reference_compare_summary.csv")
)

knitr::kable(
  pcdwba_summary,
  digits = 6,
  col.names = c(
    "Comparison",
    "Max abs. delta TS (dB)",
    "Mean abs. delta TS (dB)"
  )
)
```

The `ZooScatR` source-level reconstruction and `acousticTS` are indistinguishable on this grid. The `echopop` comparison remains close as well, but it is not at machine precision because that implementation evaluates the cylindrical Bessel term through interpolation rather than a direct nodewise call. On this grid, the largest mismatch occurs near `112 kHz`; replacing the interpolated `J_1(x)/x` evaluation with a direct call collapses that residual onto the `acousticTS` / `ZooScatR` curve. So the remaining drift is numerical, not geometrical.

## Timings

```{r echo = FALSE}
pcdwba_timing <- utils::read.csv(
  file.path("..", "..", "scratch", "pcdwba_reference_compare_timing.csv")
)

knitr::kable(
  pcdwba_timing,
  digits = 4,
  col.names = c(
    "Implementation",
    "Elapsed (s)",
    "f min (kHz)",
    "f max (kHz)",
    "Step (kHz)",
    "Node count"
  )
)
```

## Spectrum overlay

```{r fig.width = 8, fig.height = 6}
pcdwba_compare <- utils::read.csv(
  file.path("..", "..", "scratch", "pcdwba_reference_compare.csv")
)
pcdwba_compare <- pcdwba_compare[order(pcdwba_compare$frequency_hz), ]

par(mfrow = c(2, 1), mar = c(4, 4.5, 2, 1))

plot(
  pcdwba_compare$frequency_khz,
  pcdwba_compare$TS_zooscatr,
  type = "l",
  lwd = 2.5,
  col = "black",
  xlab = "",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2))
)
lines(pcdwba_compare$frequency_khz, pcdwba_compare$TS_acousticts, col = "#0b6e4f", lwd = 2)
lines(pcdwba_compare$frequency_khz, pcdwba_compare$TS_echopop, col = "#b2472f", lwd = 2, lty = 2)
legend(
  "bottomleft",
  legend = c("ZooScatR source", "acousticTS", "echopop"),
  col = c("black", "#0b6e4f", "#b2472f"),
  lwd = c(2.5, 2, 2),
  lty = c(1, 1, 2),
  bty = "n"
)

plot(
  pcdwba_compare$frequency_khz,
  pcdwba_compare$TS_acousticts - pcdwba_compare$TS_zooscatr,
  type = "l",
  lwd = 2,
  col = "#0b6e4f",
  xlab = "Frequency (kHz)",
  ylab = expression(Delta * TS ~ (dB)),
  ylim = range(c(
    pcdwba_compare$TS_acousticts - pcdwba_compare$TS_zooscatr,
    pcdwba_compare$TS_acousticts - pcdwba_compare$TS_echopop
  ))
)
lines(
  pcdwba_compare$frequency_khz,
  pcdwba_compare$TS_acousticts - pcdwba_compare$TS_echopop,
  col = "#b2472f",
  lwd = 2,
  lty = 2
)
abline(h = 0, lty = 3, col = "grey50")
legend(
  "topright",
  legend = c("acousticTS - ZooScatR source", "acousticTS - echopop"),
  col = c("#0b6e4f", "#b2472f"),
  lwd = 2,
  lty = c(1, 2),
  bty = "n"
)
```

# Closing note

This is the kind of implementation check that matters for a phase-compensated bent-cylinder solver. The comparison is not just against a benchmark curve. It is against two independently written source routines that share the same governing model. On this reference case, `acousticTS` reproduces the direct ZooScatR-style calculation exactly and stays very close to the `echopop` implementation across the full frequency band.
