---
title: "Building Scatterers"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Building Scatterers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

A shape records geometry. A scatterer assigns that geometry a physical meaning
by adding material properties, orientation, components, and metadata. Most
models operate on a scatterer rather than on a bare shape.

Scatterers are S4 objects derived from the [`Scatterer`
class](../../reference/Scatterer-class.html). Their subclasses represent
different interface physics and component structures. Choose a class from what
the target contains physically, not only from its outline.

::: {.tip data-title="Prepare the material inputs first"}
Use [Material Properties](../material-properties/material-properties.html) to
prepare densities, sound speeds, contrasts, and elastic properties before
attaching them to a shape.
:::

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_clickable_figure("building_scatterers_map")
```

::: {.tip data-title="Explore the class hierarchy"}
Select a class in the figure to open its reference page. `ELA` and `CSC` are
parent classes that organize related targets. They are not separate constructor
choices.
:::

# Choosing a scatterer class

| Constructor and class | Physical representation |
|---|---|
| [`fls_generate()`](../../reference/fls_generate.html), [`FLS`](../../reference/FLS-class.html) | A single fluid-like body |
| [`gas_generate()`](../../reference/gas_generate.html), [`GAS`](../../reference/GAS-class.html) | A gas-filled body or inclusion |
| [`sbf_generate()`](../../reference/sbf_generate.html), [`SBF`](../../reference/SBF-class.html) | A fish-like body and swimbladder |
| [`bbf_generate()`](../../reference/bbf_generate.html), [`BBF`](../../reference/BBF-class.html) | A fluid-like body and elastic backbone |
| [`ess_generate()`](../../reference/ess_generate.html), [`ESS`](../../reference/ESS-class.html) | An elastic shell enclosing a fluid |
| [`cal_generate()`](../../reference/cal_generate.html), [`CAL`](../../reference/CAL-class.html) | A solid calibration sphere |

The same geometry can support more than one physical representation. A sphere,
for example, can be used for a gas inclusion, an elastic shell, or a calibration
target. Its scatterer class determines which materials, interfaces, and models
are relevant.

# Single-component targets

The recommended workflow is to construct the geometry first and pass the
resulting `Shape` through `shape`. Material properties may be supplied as
absolute values or as contrasts, depending on the constructor and available
data:

```{r}
library(acousticTS)

body_shape <- prolate_spheroid(
  length_body = 0.04,
  radius_body = 0.004,
  n_segments = 60
)

fls_obj <- fls_generate(
  shape = body_shape,
  density_body = 1045,
  sound_speed_body = 1520,
  ID = "fluid-like target"
)

gas_obj <- gas_generate(
  shape = sphere(radius_body = 0.01, n_segments = 60),
  g_fluid = 0.0012,
  h_fluid = 0.22,
  ID = "gas target"
)
```

Plot the returned objects rather than assuming the stored geometry matches the
intended target:

```{r fig.width=8, fig.height=3.6, fig.cap="Two shapes after being assigned different physical target types."}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(3, 3, 2.2, 0.8))
plot(fls_obj, type = "shape", main = "FLS object")
plot(gas_obj, type = "shape", main = "GAS object")
par(old_par)
```

# Composite targets

Composite constructors keep acoustically distinct components separate. Build
each component as its own shape, then pass it through the corresponding named
argument:

```{r}
body_shape <- arbitrary(
  x_body = c(0, 0.04, 0.08, 0.12),
  zU_body = c(0, 0.003, 0.004, 0),
  zL_body = c(0, -0.003, -0.004, 0)
)

bladder_shape <- arbitrary(
  x_bladder = c(0.03, 0.06, 0.09),
  zU_bladder = c(0, 0.0016, 0),
  zL_bladder = c(0, -0.0016, 0)
)

sbf_obj <- sbf_generate(
  body_shape = body_shape,
  bladder_shape = bladder_shape,
  density_body = 1040,
  sound_speed_body = 1500,
  density_bladder = 1.2,
  sound_speed_bladder = 340,
  ID = "body and bladder"
)
```

An `SBF` retains `body` and `bladder` components. A `BBF` similarly retains a
`body` and `backbone`. The latter requires a cylindrical backbone with its own
density and longitudinal and transverse sound speeds:

```{r eval=FALSE}
bbf_obj <- bbf_generate(
  body_shape = body_shape,
  backbone_shape = cylinder(
    length_body = 0.06,
    radius_body = 0.0008,
    n_segments = 40
  ),
  density_body = 1070,
  sound_speed_body = 1570,
  density_backbone = 1900,
  sound_speed_longitudinal_backbone = 3500,
  sound_speed_transversal_backbone = 1700
)
```

Use [`offset_component()`](../../reference/offset_component.html) after
construction if an internal component needs a deliberate positional adjustment.

# Shells and calibration targets

`ESS` and `CAL` encode more specialized physics. An elastic shell requires
shell and enclosed-fluid properties. A calibration target uses a supported
material preset or explicitly supplied elastic properties:

```{r}
ess_obj <- ess_generate(
  shape = sphere(radius_body = 0.03, n_segments = 80),
  shell_thickness = 0.001,
  density_shell = 1050,
  sound_speed_shell = 2350,
  density_fluid = 1030,
  sound_speed_fluid = 1500,
  E = 3.5e9,
  nu = 0.34
)

cal_obj <- cal_generate(
  material = "WC",
  diameter = 38.1e-3,
  n_segments = 120
)
```

Consult the constructor reference pages before substituting materials or
elastic constants. These inputs define the boundary response, not merely object
metadata.

# Input conventions

The public constructors follow these conventions:

- Build geometry with a `Shape` constructor and pass the object through
  `shape`, `body_shape`, `bladder_shape`, or `backbone_shape`.
- Supply geometry in meters and orientation in radians. Older unit arguments
  remain for compatibility, but non-SI values are deprecated and ignored.
- For a given component property, use either the contrast form (`g_*`, `h_*`)
  or the absolute form (`density_*`, `sound_speed_*`).
- Use `ID` for a stable target identifier.
- Keep components separate when their interfaces matter acoustically.

Raw coordinate inputs remain available in some constructors for older code.
Character labels that ask a scatterer constructor to generate a shape are also
retained for compatibility. New workflows should use pre-built `Shape` objects
so geometry can be inspected before physical properties are added.

# Check the object

Use [`extract()`](../../reference/extract.html) and
[`plot()`](../../reference/plot.Scatterer.html) to verify the object before
running a model:

```{r}
class(sbf_obj)
names(extract(sbf_obj, "body"))
names(extract(sbf_obj, "bladder"))
extract(sbf_obj, "metadata")
```

Confirm the following:

- the S4 class matches the intended physical target
- component geometry and relative placement are plausible
- materials or contrasts use the intended reference medium
- orientation and units are correct
- all interfaces needed by the model remain distinct

Continue with [Running Models](../running-models/running-models.html) and
[Choosing a Model](../model-selection/model-selection.html).
