---
title: "Shape Manipulation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Shape Manipulation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

Geometry can be repositioned, bent, resized, or resampled without rebuilding a
target from its original inputs. These operations are useful for cleaning
measured profiles and for controlled studies of posture, size, component
placement, and numerical resolution.

Manipulation changes the object returned by the package. It can also change the
scientific meaning of that object. Keep the original object, give each modified
variant a distinct name, and record what the transformation represents.

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_clickable_figure("shape_manipulation_schematic")
```

::: {.tip data-title="Open the main function references"}
Select [`brake()`](../../reference/brake.html) or
[`reforge()`](../../reference/reforge.html) in the figure to open its reference
page.
:::

# Choose the operation

| Function | Use |
|---|---|
| [`brake()`](../../reference/brake.html) | Bend a centerline using relative or measured curvature. |
| [`reforge()`](../../reference/reforge.html) | Resize components, set target dimensions, or change segment counts. |
| [`translate_shape()`](../../reference/translate_shape.html) | Apply coordinate offsets. |
| [`reanchor_shape()`](../../reference/reanchor_shape.html) | Place the nose, center, or tail at a requested coordinate. |
| [`offset_component()`](../../reference/offset_component.html) | Move one component within a composite scatterer. |
| [`inflate_shape()`](../../reference/inflate_shape.html) | Widen or narrow a selected axial region. |
| [`smooth_shape()`](../../reference/smooth_shape.html) | Smooth a stored profile. |
| [`resample_shape()`](../../reference/resample_shape.html) | Change axial discretization. |
| [`flip_shape()`](../../reference/flip_shape.html) | Reverse the axial profile. |

Use `brake()` for posture or curvature. Use `reforge()` for size, proportions,
or resolution. The smaller helpers handle positioning and profile preparation.

# Bending with `brake()`

`brake()` applies a smooth curvature to a body component or supported scatterer.
In `mode = "ratio"`, `radius_curvature` is relative to body length. In
`mode = "measurement"`, it is a physical radius in the geometry's length
units. Ratio mode supports size-normalized comparisons. Measurement mode is
appropriate when curvature has been measured directly.

```{r}
library(acousticTS)

straight_obj <- fls_generate(
  shape = cylinder(
    length_body = 0.05,
    radius_body = 0.003,
    n_segments = 120
  ),
  density_body = 1045,
  sound_speed_body = 1520
)

bent_ratio <- brake(
  straight_obj,
  radius_curvature = 5,
  mode = "ratio"
)

bent_measured <- brake(
  straight_obj,
  radius_curvature = 0.35,
  mode = "measurement"
)
```

```{r fig.width=8, fig.height=3.6, fig.cap="A fluid-like cylinder before and after ratio-based bending."}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(3.2, 3.2, 2.6, 0.8))
plot(straight_obj, type = "shape", main = "Original FLS cylinder")
plot(bent_ratio, type = "shape", main = "After brake()")
par(old_par)
```

Curvature changes the projected length, local orientation, and phase
relationships along the body. Check the modified centerline and segment
resolution, not only the outline.

::: {.warning data-title="Length of a bent FLS object"}
For a bent `FLS`, stored body length is the centerline arc length rather than
the projected `x` span. If `reforge()` is applied afterward, a requested length
rescales that bent centerline. To resize a straight target and bend it afterward,
apply the functions in that order explicitly.
:::

The distinction can be checked directly:

```{r}
bent_rescaled <- reforge(
  bent_ratio,
  body_target = c(length = 0.08)
)

bent_rpos <- extract(bent_rescaled, "body")$rpos
centerline_length <- sum(sqrt(
  diff(bent_rpos["x", ])^2 + diff(bent_rpos["z", ])^2
))

c(
  stored_length = extract(
    bent_rescaled,
    c("shape_parameters", "length")
  ),
  centerline_length = centerline_length,
  projected_x_span = diff(range(bent_rpos["x", ]))
)
```

# Resizing with `reforge()`

`reforge()` is an S4 generic with methods for `FLS`, `GAS`, `SBF`, and `BBF`
objects. Its arguments depend on the class, but two forms recur:

- `*_scale` multiplies current dimensions.
- `*_target` requests final dimensions.

Use named dimensions such as `length`, `width`, `height`, and `radius`. Set the
corresponding `isometric_*` argument to `FALSE` when the dimensions should vary
independently. Segment-count arguments change numerical resolution without
requiring a separate reconstruction.

```{r}
data(krill, package = "acousticTS")

krill_40mm <- reforge(
  krill,
  body_target = c(length = 0.04)
)

krill_fine <- reforge(
  krill,
  n_segments_body = 200
)

data.frame(
  object = c("original", "40 mm", "200 segments"),
  length_m = c(
    extract(krill, c("shape_parameters", "length")),
    extract(krill_40mm, c("shape_parameters", "length")),
    extract(krill_fine, c("shape_parameters", "length"))
  ),
  n_segments = c(
    extract(krill, c("shape_parameters", "n_segments")),
    extract(krill_40mm, c("shape_parameters", "n_segments")),
    extract(krill_fine, c("shape_parameters", "n_segments"))
  )
)
```

## Composite scatterers

For `SBF` and `BBF` objects, body and internal components can be changed
separately. The following example lengthens the body, reduces swimbladder
height, and resamples both components:

```{r}
data(sardine, package = "acousticTS")

sardine_modified <- reforge(
  sardine,
  body_scale = c(length = 1.2),
  swimbladder_scale = c(height = 0.8),
  isometric_body = FALSE,
  isometric_swimbladder = FALSE,
  maintain_ratio = FALSE,
  n_segments_body = 60,
  n_segments_swimbladder = 40,
  containment = "warn"
)
```

Height changes are applied about each component's own vertical center. For
upper/lower envelope profiles, this is the implied local midline. For explicit
centerlines, such as an `FLS` body or `BBF` backbone, the path remains fixed
while the surrounding thickness changes.

After changing an internal component, verify that it remains inside the body.
The `containment` argument can warn, stop with an error, or skip this check. Use
[`offset_component()`](../../reference/offset_component.html) when a bladder or
backbone needs a controlled positional adjustment.

## Canonical gas-filled bodies

For canonical `GAS` bodies, `reforge()` regenerates the geometry with the
matching shape constructor instead of scaling the stored coordinates directly.
This preserves a clean discretization and an accurate shape label.

```{r}
gas_sphere <- gas_generate(
  shape = sphere(radius_body = 0.01, n_segments = 40),
  g_fluid = 0.0012,
  h_fluid = 0.22
)

gas_isometric <- reforge(
  gas_sphere,
  body_target = c(radius = 0.015)
)

gas_elongated <- reforge(
  gas_sphere,
  body_target = c(length = 0.05),
  isometric_body = FALSE
)

c(
  original = extract(gas_sphere, c("shape_parameters", "shape")),
  isometric = extract(gas_isometric, c("shape_parameters", "shape")),
  elongated = extract(gas_elongated, c("shape_parameters", "shape"))
)
```

A sphere remains spherical only when length and radius retain the spherical
relationship. An independent length change promotes it to a prolate spheroid.
Arbitrary `GAS` bodies have no canonical constructor to regenerate them, so
they are scaled directly by axis.

# Positioning and profile preparation

The smaller manipulation helpers can be combined for measured profiles:

```{r}
centered_obj <- reanchor_shape(
  straight_obj,
  anchor = "center",
  at = 0
)

prepared_obj <- straight_obj |>
  inflate_shape(x_range = c(0.015, 0.035), scale = 1.35) |>
  smooth_shape(span = 7) |>
  resample_shape(n_segments = 120)

reversed_obj <- flip_shape(prepared_obj, axis = "x")
```

Profile-changing operations can convert canonical geometry into an arbitrary
representation. Recheck the stored shape label and model compatibility after
local edits. Translation and reanchoring change position but not dimensions.

# Checks after manipulation

Before running another model:

- plot the original and modified objects together
- inspect component lengths, radii, and relative positions
- confirm the stored shape class or label
- check segment convergence after resampling
- verify containment for composite targets
- record whether the change represents biology, posture, or numerical cleanup
- rerun the model rather than reusing results stored before the transformation

Continue with [Building Shapes](../building-shapes/building-shapes.html),
[Building Scatterers](../building-scatterers/building-scatterers.html), or
[Comparing Models](../comparing-models/comparing-models.html).
