---
title: "Target strength for a calibration sphere"
author: "Brandyn Lucca (https://orcid.org/0000-0003-3145-2969)"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Target strength for a calibration sphere}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction 

Echosounders are often calibrated using standard targets that comprise strong
scatterers with target strengths (TS, dB re. 1 m^2^) that are relatively 
straightforward to model and predict[^1]. Typically, this comprises
a tungsten carbide (chemical formula: WC) sphere, but elastic spheres consisting
of other materials (e.g. aluminum, Al) can also be used[^2]. 

[^1]: Foote, K.G. (**1990**). *Spheres for calibration an eleven-frequency acoustic
measurement system*. J. Cons. int. Explor. Mer., 46: 284-286. 

[^2]: Hickling, R. (**1962**). *Analysis of echoes from a solid elastic sphere in water*.
J. Acoust. Soc. Am., 45: 1582-1592. 

# acousticTS implementation

The `acousticTS` package adapts the series of equations described in the 
literature[^3] that provide a solution of wave equations based on the classical
theory of elasticity. The object-based approach in `acousticTS` makes this 
relatively straightforward by minimizing the amount of manual coding end-users
must write. 

[^3]: MacLennan, D.N. (**1981**). *Target strength measurements on metal spheres*. 
Scottis Fish. Res. Rep., 25, 11 pp. 

## Calibration sphere object generation

First, a calibration sphere object has to be created. This is represented by the
`CAL` object class that contains slots for `metadata`, `model_parameters`, 
`model` results, `body` for sphere dimensions and material properties, and
`shape_parameters` for shape-specific metadata. This object can be created using
the `cal_generate(...)` function that has two required arguments: `material` and 
`diameter`. The default diameter is 38.1 mm, or 38.1e^-3^ m. The `diameter` input
is intended to be in meters.The `material` argument comprises five default 
options that automatically include their respective longitudinal and transversal 
sound speeds (m s^-1^) and material density (kg m^-3^): 


| Material | Argument value | Longitudinal sound speed (c~l~) | Transversal sound speed (c~t~) | Density ($\rho$) |
| ----------- | ----------- | ----------- | ----------- | ----------- |
| Tungsten carbide | "WC" | 6853 | 4171 | 8360 |
| Aluminum | "Al" | 6260 | 3080 | 2070 |
| Stainless steel | "steel" | 5980 | 3297 | 7970 |
| Brass | "brass" | 4372 | 2100 | 8360 |
| Copper | "Cu" | 4760 | 2288.5 |8947 |

Alternatively, you can define your own material properties by assigning values to 
`sound_speed_longitudinal`, `sound_speed_transversal`, and `density_sphere` within
`cal_generate(...)` such as: 

`cal_generate(..., density_sphere = 1026)`

When using the default arguments: 

```{r}
# Call in package library
library(acousticTS)
# Create calibration sphere object
cal_sphere <- cal_generate()
# this is equivalent to: cal_sphere <- cal_generate(material = "WC", diameter = 38.1e-3)
```

## Calculating a TS-frequency spectrum for the calibration sphere

With the calibration sphere object generated, TS can be calculated via the 
`target_strength(...)` function, which is a wrapper function that generally
allows for multiple models to be applied to a single scatterer when needed. In 
this case, there are three required arguments to calculate TS: `object`, 
`frequency`, and `model`. The `object` argument simply refers to the `CAL` object
that we created before (i.e. `cal_sphere`). Frequency (via `frequency`) can be 
a `vector` of values (Hz). Model is a `string` input that refers to the model 
end-users would like to use, which would be `model = "calibration"` in this case.
This will update the current `CAL` object when assigned to the same variable, 
or can be used to generate a copy of the original `CAL` object that now includes
model parameter and results as a built-in field.


```{r} 
# Define frequency vector 
frequency <- seq(1e3, 600e3, 1e3)
# Calculate TS; update original CAL object
cal_sphere <- target_strength(object = cal_sphere,
                              frequency = frequency, 
                              model = "calibration")
# Calculate TS; store in a new CAL object
cal_sphere_copy <- target_strength(object = cal_sphere,
                                   frequency = frequency, 
                                   model = "calibration")
```

## Extracting model results

Model results can be extracted either just visually, or can be directly accessed
using the `extract(...)` function. 

### Plotting results

This approach uses the `plot(...)` generic function with additional arguments
for toggling the plot output. The additional arguments to include here are:

- `type = "shape"` *or* `type = "model"`
- `nudge_y = 1.05` *(default)*
- `nudge_x = 1.01` *(default)*
- `x_units = "frequency"` *or* `x_units = "k_sw"` *or* 
`x_units = "k_l"` *or* `x_units = "k_t"`

Setting `type = "model"` will parse the model results (i.e. TS). The two `nudge_` 
arguments allow the end-user to nudge to *x*- and *y*-axes as they see fit. The
`x_units` argument defaults to `x_units = "frequency"`, but setting it to equal 
`"k_sw"`, `"k_l"`, or `"k_t"` will set TS to be a function of the wavenumber of 
seawater, longitudinal axis of sphere, and transversal axis of sphere, 
respectively multiplied by the sphere's radius. 

```{r fig.width = 6}
# Plot TS as a function of frequency (kHz)
plot(cal_sphere, type = "model", ylim = c(-70, -30))
# Plot TS as a function of k[sw]*radius, or k[sw]*a
plot(cal_sphere, type = "model", x_units = "k_sw", ylim = c(-70, -30))
# Plot TS as a function of k[l]*radius, or k[sw]*a
plot(cal_sphere, type = "model", x_units = "k_l", ylim = c(-70, -30))
# Plot TS as a function of k[t]*radius, or k[sw]*a
plot(cal_sphere, type = "model", x_units = "k_t", ylim = c(-70, -30))
# Nudge the x-axis more
plot(cal_sphere, type = "model", ylim = c(-70, -30))
# Nudge the y-axis more
plot(cal_sphere, type = "model", ylim = c(-70, -30))
```

### Accessing results

The model results can also be directly accessed via `extract(...)` which has the
two arguments `object` and `feature`. In this case, `object` refers to our `CAL`
object, `cal_sphere`, and we can set `feature = "model"` to extract our model
results. 

```{r}
# Extract model results
model_results <- extract( cal_sphere, "model" )$calibration
# Peek at the output from extract(...)
head( model_results )

```

Here we get seven columns formatted as a `data.frame` object: 

- `frequency`: transmit frequency (Hz)
- `k_sw`: acoustic wavenumber for seawater/ambient fluid
- `k_l`: longitudinal axis acoustic wavenumber of sphere
- `k_t`: transversal axis acoustic wavenumber of sphere
- `f_bs`: the complex form function output
- `sigma_bs`: the acoustic cross-section for an elastic sphere (m^2^) where: 
$\sigma_{bs} = \pi a^2 ~|f_{bs}|^2$
- `TS`: the target strength (dB re. 1 m^2^) for an elastic sphere where: 
$TS = 10log_{10}(\frac{\sigma_{bs}}{4 \pi})$

# Material property and sphere shape comparisons

This implementation also allows for comparisons among different model parameters
and diameters suited for specific calibration experiments. 

## Comparing different diameters
```{r fig.width = 8, fig.height = 4.5}
# Generate models for calibration spheres
# 20 mm diameter sphere
sphere_20 <- cal_generate(diameter = 20e-3)
sphere_20 <- target_strength(object = sphere_20,
                             frequency = frequency,
                             model = "calibration")
ts_mods_20 <- extract(sphere_20, "model")$calibration
# 30 mm diameter sphere
sphere_30 <- cal_generate(diameter = 30e-3)
sphere_30 <- target_strength(object = sphere_30,
                             frequency = frequency,
                             model = "calibration")
ts_mods_30 <- extract(sphere_30, "model")$calibration
# 40 mm diameter sphere
sphere_40 <- cal_generate(diameter = 40e-3)
sphere_40 <- target_strength(object = sphere_40,
                             frequency = frequency,
                             model = "calibration")
ts_mods_40 <- extract(sphere_40, "model")$calibration
# Plot the three and compare
par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_mods_20$frequency*1e-3, 
     y = ts_mods_20$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2,
     ylim = c(-70, -30))
lines(x = ts_mods_30$frequency*1e-3,
      y = ts_mods_30$TS,
      col = 'red',
      lty = 1,
      lwd = 2.5)
lines(x = ts_mods_40$frequency*1e-3,
      y = ts_mods_40$TS,
      col = 'blue',
      lty = 1,
      lwd = 2.5)
legend("bottomright",
       c("20 mm","30 mm", "40 mm"),
       lty=c(1, 1, 1),
       lwd=c(2.5, 2.5, 2.5),
       col=c('black', 'red', 'blue'),
       cex=1.1)

```

## Comparing different materials

```{r fig.width = 8, fig.height = 4.5}
# tungsten carbide
ts_mods_WC <- extract(cal_sphere, "model")$calibration
# copper sphere
sphere_copper <- cal_generate(material = "Cu")
sphere_copper <- target_strength(object = sphere_copper,
                                 frequency = frequency,
                                 model = "calibration")
ts_mods_Cu <- extract(sphere_copper, "model")$calibration
# aluminum sphere
sphere_aluminum <- cal_generate(material = "Al")
sphere_aluminum <- target_strength(object = sphere_aluminum,
                                   frequency = frequency,
                                   model = "calibration")
ts_mods_Al <- extract(sphere_aluminum, "model")$calibration
# brass sphere
sphere_brass <- cal_generate(material = "brass")
sphere_brass <- target_strength(object = sphere_brass,
                                frequency = frequency,
                                model = "calibration")
ts_mods_brass <- extract(sphere_brass, "model")$calibration
# stainless steel sphere
sphere_steel <- cal_generate(material = "steel")
sphere_steel <- target_strength(object = sphere_steel,
                                frequency = frequency,
                                model = "calibration")
ts_mods_steel<- extract(sphere_steel, "model")$calibration
# Plot each and compare
par(ask = F,
    oma = c(1, 1, 1, 0),
    mar = c(3, 4.5, 1, 2))
plot(x = ts_mods_WC$frequency*1e-3, 
     y = ts_mods_WC$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2,
     ylim = c(-90, -30))
lines(x = ts_mods_Cu$frequency*1e-3,
      y = ts_mods_Cu$TS,
      col = 'red',
      lty = 1,
      lwd = 2.5)
lines(x = ts_mods_Al$frequency*1e-3,
      y = ts_mods_Al$TS,
      col = 'blue',
      lty = 1,
      lwd = 2.5)
lines(x = ts_mods_brass$frequency*1e-3,
      y = ts_mods_brass$TS,
      col = 'green',
      lty = 1,
      lwd = 2.5)
lines(x = ts_mods_steel$frequency*1e-3,
      y = ts_mods_steel$TS,
      col = 'orange',
      lty = 1,
      lwd = 2.5)
legend("bottomright",
       c("WC","Cu", "Al", "Brass", "Steel"),
       lty=c(1, 1, 1, 1, 1),
       lwd=c(2.5, 2.5, 2.5, 2.5, 2.5),
       col=c('black', 'red', 'blue', 'green', 'orange'),
       cex=1.1)  

```

# Future development

Stay tuned for new methods enabling backscatter simulations for calibration 
spheres of different materials, sizes, etc., from a single function rather than 
repeating the same dataflow.
