---
title: "DWBA and SDWBA models for fluid-like scatterers"
author: "Brandyn Lucca (https://orcid.org/0000-0003-3145-2969)"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{DWBA and SDWBA models for fluid-like scatterers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction 

The Distorted-Wave Born Approximation (DWBA) is a widely used theoretical framework
for modeling acoustic backscatter from fluid-like organisms such as zooplankton[^1][^2]. 
The DWBA treats scatterers as weakly-scattering objects with material properties 
that differ only slightly from the surrounding seawater. This approximation is 
particularly well-suited for modeling crustacean zooplankton like copepods and krill, 
which have body densities and sound speeds close to those of seawater.

[^1]: Stanton, T.K., Chu, D., and Wiebe, P.H. (1998). *Sound scattering by several
zooplankton groups. II. Scattering models*. J. Acoust. Soc. Am., 103, 236-253.

[^2]: Demer, D.A., and Conti, S.G. (2003). *Reconciling theoretical versus empirical
target strengths of krill: effects of phase variability on the distorted-wave
Born approximation*. ICES J. Mar. Sci., 60: 429-434.

# acousticTS implementation

The `acousticTS` package provides four DWBA implementations for fluid-like scatterers:

1. **DWBA**: Standard DWBA for straight-bodied organisms
2. **DWBA_curved**: DWBA for organisms with curved body shapes
3. **SDWBA**: Stochastic DWBA accounting for phase variability in straight bodies
4. **SDWBA_curved**: Stochastic DWBA for curved organisms with phase variability

Both deterministic (DWBA) and stochastic (SDWBA) models are applied to objects of 
the `FLS` (Fluid-Like Scatterer) class, which can represent various zooplankton 
shapes from simple cylinders to complex curved forms.

```{r}
# Call in package library
library(acousticTS)

# Create a simple cylinder shape
cylinder_shape <- cylinder(
  length_body = 15e-3,     # 15 mm length
  radius_body = 2e-3,      # 2 mm radius
  n_segments = 50          # 50 discrete segments
)

# Create FLS object with the cylinder shape
cylinder_scatterer <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.058,          # Density contrast (ρ_body/ρ_water)
  h_body = 1.058,          # Sound speed contrast (c_body/c_water)
  theta_body = pi/2        # Broadside orientation
)

# Display the object
cylinder_scatterer
```

## Using the krill dataset

The `acousticTS` package includes a built-in krill dataset based on the digitized
body shape from McGehee et al. (1998)[^3], which provides realistic morphological
and material properties for Antarctic krill:

[^3]: McGehee, D.E., O'Driscoll, R.L., and Traykovski, L.V.M. (1998). *Effects of 
orientation on acoustic scattering from Antarctic krill at 120 kHz*. Deep Sea Res. 
Part II: Top. Stud. Oceanogr. 45, 1273–1294.

```{r}
# Load the krill dataset
data(krill)

# Display basic information about the krill object
krill
```

Let's examine the krill shape:

```{r fig.width = 8, fig.height = 5}
# Plot the krill shape
plot(krill, type = "shape")
```

# DWBA model calculations

## Standard DWBA for cylindrical scatterer

The standard DWBA model assumes straight-bodied organisms and is appropriate for 
simple cylindrical shapes:

```{r}
# Define frequency vector 
frequency <- seq(50e3, 500e3, 10e3)  # 50 kHz to 500 kHz

# Calculate TS using standard DWBA
cylinder_scatterer <- target_strength(
  object = cylinder_scatterer,
  frequency = frequency,
  model = "DWBA"
)
```

## DWBA models for krill

For the krill dataset, we can apply all four DWBA models to compare their performance.
The curved models are more appropriate for krill due to their naturally curved body shape,
while the stochastic models account for phase variability:

```{r}
# Define radius-of-curvature for curved models
krill@body$radius_curvature_ratio <- 3.0

# Apply all DWBA models to krill
krill <- target_strength(
  object = krill,
  frequency = frequency,
  model = c("DWBA", "DWBA_curved", "SDWBA", "SDWBA_curved")
)
```

# Visualizing results

## Cylindrical scatterer results

```{r fig.width = 6, fig.height = 4}
# Plot TS for the cylindrical scatterer
plot(cylinder_scatterer, type = "model")
```

## Krill model comparison

```{r fig.width = 8, fig.height = 5}
# Plot all DWBA models for krill
plot(krill, type = "model")
```

# Model parameters and sensitivity

## Extracting model results

Model results contain detailed information about the scattering calculations:

```{r}
# Extract DWBA results for krill
dwba_results <- extract(krill, "model")$DWBA
head(dwba_results)

# Extract curved DWBA results
dwba_curved_results <- extract(krill, "model")$DWBA_curved
head(dwba_curved_results)
```

The DWBA results include:
- `frequency`: transmit frequency (Hz)
- `k_sw`: acoustic wavenumber for seawater
- `f_bs`: complex backscattering amplitude
- `sigma_bs`: backscattering cross-section (m²)
- `TS`: target strength (dB re. 1 m²)

## Material property effects

Let's explore how different material properties affect the target strength:

```{r fig.width = 8, fig.height = 6}
# Create scatterers with different density contrasts
high_contrast <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.10,    # Higher density contrast
  h_body = 1.058,
  theta_body = pi/2
)

low_contrast <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.02,    # Lower density contrast  
  h_body = 1.058,
  theta_body = pi/2
)

# Calculate TS for both
high_contrast <- target_strength(high_contrast, frequency, "DWBA")
low_contrast <- target_strength(low_contrast, frequency, "DWBA")

# Extract results
ts_high <- extract(high_contrast, "model")$DWBA
ts_low <- extract(low_contrast, "model")$DWBA
ts_original <- extract(cylinder_scatterer, "model")$DWBA

# Plot comparison
par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_original$frequency * 1e-3, 
     y = ts_original$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2,
     ylim=c(-95, -50))

lines(x = ts_high$frequency * 1e-3,
      y = ts_high$TS,
      col = 'darkgoldenrod3',
      lty = 1,
      lwd = 2.5)

lines(x = ts_low$frequency * 1e-3,
      y = ts_low$TS,
      col = 'cornflowerblue',
      lty = 1,
      lwd = 2.5)

legend("bottomright",
       c("g = 1.058 (original)", "g = 1.10 (high)", "g = 1.02 (low)"),
       lty = c(1, 1, 1),
       lwd = c(2.5, 2.5, 2.5),
       col = c('black', 'darkgoldenrod3', 'cornflowerblue'),
       cex = 1.0)
```

## Orientation effects

The DWBA is highly sensitive to the orientation of the scatterer relative to the
incident sound wave. Let's examine this effect:

```{r fig.width = 8, fig.height = 6}
# Create scatterers at different orientations
broadside <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.058, h_body = 1.058,
  theta_body = pi/2    # 90° - broadside
)

oblique <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.058, h_body = 1.058,
  theta_body = pi/4    # 45° - oblique
)

end_on <- fls_generate(
  shape = cylinder_shape,
  g_body = 1.058, h_body = 1.058,
  theta_body = 0       # 0° - end-on
)

# Calculate TS for all orientations
broadside <- target_strength(broadside, frequency, "DWBA")
oblique <- target_strength(oblique, frequency, "DWBA")
end_on <- target_strength(end_on, frequency, "DWBA")

# Extract results
ts_broadside <- extract(broadside, "model")$DWBA
ts_oblique <- extract(oblique, "model")$DWBA
ts_end_on <- extract(end_on, "model")$DWBA

# Plot orientation comparison
par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_broadside$frequency * 1e-3, 
     y = ts_broadside$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2,
     ylim=c(-120, -55))

lines(x = ts_oblique$frequency * 1e-3,
      y = ts_oblique$TS,
      col = 'darkgoldenrod2',
      lty = 1,
      lwd = 2.5)

lines(x = ts_end_on$frequency * 1e-3,
      y = ts_end_on$TS,
      col = 'firebrick3',
      lty = 1,
      lwd = 2.5)

legend("bottomright",
       c("Broadside (90°)", "Oblique (45°)", "End-on (0°)"),
       lty = c(1, 1, 1),
       lwd = c(2.5, 2.5, 2.5),
       col = c('black', 'darkgoldenrod2', 'firebrick3'),
       cex = 1.0)
```

# Creating custom shapes

The `acousticTS` package allows for creating custom shapes beyond simple cylinders.
Here's an example using a polynomial-deformed cylinder based on Smith et al. (2013)[^4]:

[^4]: Smith, J.N., Ressler, P.H., and Warren, J.D. (2013). *A distorted wave Born
approximation target strength model for Bering Sea euphausiids*. ICES Journal
of Marine Science, 70(1): 204-214.

```{r}
# Create a polynomial deformation vector (example coefficients)
poly_coeffs <- c(0.83, 0.36, -2.10, -1.20, 0.63, 0.82, 0.64)  # Smith et al. (2013)

# Create a polynomial-deformed cylinder
poly_shape <- polynomial_cylinder(
  length_body = 15e-3,
  radius_body = 2e-3,
  n_segments = 50,
  polynomial = poly_coeffs
)

# Create FLS object with polynomial shape
poly_scatterer <- fls_generate(
  shape = poly_shape,
  g_body = 1.058,
  h_body = 1.058,
  radius_curvature_ratio = 3.0,
  theta_body = pi/2
)

# Calculate TS
poly_scatterer <- target_strength(
  object = poly_scatterer,
  frequency = frequency,
  model = "DWBA_curved"  # Use curved DWBA for deformed shapes
)
```

```{r fig.width = 8, fig.height = 5}
# Plot the polynomial-deformed shape
plot(poly_scatterer, type = "shape")
```

```{r fig.width = 8, fig.height = 5}
# Compare with regular cylinder
ts_poly <- extract(poly_scatterer, "model")$DWBA_curved
ts_cylinder <- extract(cylinder_scatterer, "model")$DWBA

par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_cylinder$frequency * 1e-3, 
     y = ts_cylinder$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     col = 'cornflowerblue',
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2,
     ylim=c(-100, -55))

lines(x = ts_poly$frequency * 1e-3,
      y = ts_poly$TS,
      col = 'darkseagreen4',
      lty = 2,
      lwd = 2.5)

legend("bottomright",
       c("Regular cylinder", "Polynomial-deformed cylinder"),
       lty = c(1, 2),
       lwd = c(2.5, 2.5),
       col = c('cornflowerblue', 'darkseagreen4'),
       cex = 1.0)
```

# Model applications and validity

## When to use DWBA vs. DWBA_curved

- **DWBA**: Best for straight-bodied organisms or when computational speed is important
- **DWBA_curved**: More accurate for naturally curved organisms like krill, but computationally more intensive

## Biological applications

The DWBA models are particularly well-suited for:

- **Copepods**: Small crustacean zooplankton with approximately cylindrical bodies
- **Krill**: Larger crustaceans with naturally curved body shapes
- **Chaetognaths**: Arrow worms with elongated, transparent bodies
- **Small fish larvae**: When swim bladders are not yet developed

## Model limitations

The DWBA has several important limitations:

1. **Weak scattering assumption**: Valid only when material contrasts are small (g, h ≈ 1)
2. **Single scattering**: Does not account for multiple scattering within the organism
3. **Shape assumptions**: Assumes cylindrical segments; complex 3D shapes may require other approaches

# Future development

Future enhancements may include:

- Stochastic DWBA implementations (SDWBA)
- Integration with empirical orientation distributions
- Automated parameter estimation from morphological data
- Support for temperature and pressure-dependent material properties
