---
title: "Target strength for elastic shelled spheres"
author: "Brandyn Lucca (https://orcid.org/0000-0003-3145-2969)"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Target strength for elastic shelled spheres}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction 

Elastic shelled spheres are important scatterers in marine acoustics, representing
various zooplankton species and other organisms with hard exoskeletons or shells.
These scatterers consist of an outer elastic shell surrounding an inner fluid
medium. The target strength (TS, dB re. 1 m^2^) of such scatterers can be modeled
using either exact modal series solutions[^1] or high-frequency approximations[^2].

[^1]: Goodman, R.R., and Stern, R. (1962). *Reflection and transmission of sound by 
elastic spherical shells*. The Journal of the Acoustical Society of America,
34, 338-344.

[^2]: Stanton, T.K. (1989). Simple approximate formulas for backscattering of sound by spherical and elongated objects. The Journal of the Acoustical Society of America, 86, 1499-1510.

# acousticTS implementation

The `acousticTS` package provides two main approaches for modeling elastic shelled
spheres: the exact modal series solution from Goodman and Stern (1962) and the
high-frequency ray-based approximation from Stanton (1989). The object-based
approach in `acousticTS` makes implementing these models straightforward by
handling the complex mathematical calculations internally.

## Elastic shelled sphere object generation

First, an elastic shelled sphere object must be created using the `ESS` (Elastic
Shelled Scatterer) object class. This contains slots for `metadata`, 
`model_parameters`, `model` results, `shell` properties, `fluid` properties,
and `shape_parameters`. The object can be created using the `ess_generate(...)`
function.

### Required parameters

The `ess_generate(...)` function requires several key parameters:

- `radius_shell`: Outer radius of the shell (m)
- `shell_thickness`: Thickness of the shell (m) [optional, can be calculated from inner/outer radii]
- Shell material properties: density, sound speed, and elastic constants
- Fluid material properties: density and sound speed

### Material properties for the shell

The shell requires specification of:
- `density_shell` or `g_shell`: Shell density (kg/m³) or density contrast
- `sound_speed_shell` or `h_shell`: Shell sound speed (m/s) or sound speed contrast
- Elastic constants: `K` (bulk modulus), `nu` (Poisson's ratio), `G` (shear modulus), or `E` (Young's modulus)

### Material properties for the internal fluid

The internal fluid requires:
- `density_fluid` or `g_fluid`: Fluid density (kg/m³) or density contrast
- `sound_speed_fluid` or `h_fluid`: Fluid sound speed (m/s) or sound speed contrast

### Basic example

```{r}
# Call in package library
library(acousticTS)

# Create elastic shelled sphere object
shelled_sphere <- ess_generate(
  shape = "sphere",
  radius_shell = 10e-3,        # 10 mm outer radius
  shell_thickness = 1e-3,      # 1 mm shell thickness
  sound_speed_shell = 3750,    # Shell sound speed (m/s)
  sound_speed_fluid = 1575,    # Internal fluid sound speed (m/s)
  density_shell = 2565,        # Shell density (kg/m³)
  density_fluid = 1077.3,      # Internal fluid density (kg/m³)
  K = 70e9,                    # Bulk modulus (Pa)
  nu = 0.32                    # Poisson's ratio
)

# Display the object
shelled_sphere
```

## Calculating a TS-frequency spectrum for the elastic shelled sphere

With the elastic shelled sphere object generated, TS can be calculated using the
`target_strength(...)` function. This function supports multiple models for
elastic shelled spheres:

1. `"MSS_goodman_stern"`: Exact modal series solution[^1]
2. `"high_pass_stanton"`: High-frequency ray-based approximation[^2]

### Single model calculation

```{r}
# Define frequency vector 
frequency <- seq(1e3, 500e3, 1e3)  # 1 kHz to 500 kHz

# Calculate TS using the modal series solution
shelled_sphere <- target_strength(
  object = shelled_sphere,
  frequency = frequency, 
  model = "MSS_goodman_stern"
)
```

### Multiple model comparison

```{r}
# Calculate TS using both available models
shelled_sphere <- target_strength(
  object = shelled_sphere,
  frequency = frequency,
  model = c("MSS_goodman_stern", "high_pass_stanton")
)
```

## Visualizing results

### Shape visualization

The shape of the elastic shelled sphere can be visualized using:

```{r fig.width = 5.25, fig.height = 5.25}
# Plot the shape
plot(shelled_sphere, type = "shape")
```

### Model results visualization

Model results can be plotted to compare the two approaches:

```{r fig.width = 8, fig.height = 6}
# Plot TS as a function of frequency
plot(shelled_sphere, type = "model")
```

## Extracting model results

Model results can be extracted using the `extract(...)` function for further analysis:

```{r}
# Extract modal series solution results
mss_results <- extract(shelled_sphere, "model")$MSS_goodman_stern
head(mss_results)

# Extract high-pass approximation results
hp_results <- extract(shelled_sphere, "model")$high_pass_stanton
head(hp_results)
```

The modal series solution returns a `data.frame` with columns:
- `frequency`: transmit frequency (Hz)
- `k_sw`: acoustic wavenumber for seawater/ambient fluid
- `k_shell_l`: longitudinal acoustic wavenumber of shell
- `k_shell_t`: transversal acoustic wavenumber of shell
- `f_bs`: the complex form function output
- `sigma_bs`: the acoustic cross-section (m²)
- `TS`: the target strength (dB re. 1 m²)

# Parameter sensitivity analysis

This implementation allows for exploration of how different material properties
and geometric parameters affect the target strength.

## Effect of shell thickness

```{r fig.width = 8, fig.height = 4.5}
# Create spheres with different shell thicknesses
sphere_thin <- ess_generate(
  radius_shell = 10e-3, shell_thickness = 0.5e-3,
  sound_speed_shell = 3750, sound_speed_fluid = 1575,
  density_shell = 2565, density_fluid = 1077.3,
  K = 70e9, nu = 0.32
)

sphere_medium <- ess_generate(
  radius_shell = 10e-3, shell_thickness = 1.0e-3,
  sound_speed_shell = 3750, sound_speed_fluid = 1575,
  density_shell = 2565, density_fluid = 1077.3,
  K = 70e9, nu = 0.32
)

sphere_thick <- ess_generate(
  radius_shell = 10e-3, shell_thickness = 2.0e-3,
  sound_speed_shell = 3750, sound_speed_fluid = 1575,
  density_shell = 2565, density_fluid = 1077.3,
  K = 70e9, nu = 0.32
)

# Calculate TS for each sphere
sphere_thin <- target_strength(sphere_thin, frequency, "MSS_goodman_stern")
sphere_medium <- target_strength(sphere_medium, frequency, "MSS_goodman_stern")
sphere_thick <- target_strength(sphere_thick, frequency, "MSS_goodman_stern")

# Extract results
ts_thin <- extract(sphere_thin, "model")$MSS_goodman_stern
ts_medium <- extract(sphere_medium, "model")$MSS_goodman_stern
ts_thick <- extract(sphere_thick, "model")$MSS_goodman_stern

# Plot comparison
par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_thin$frequency * 1e-3, 
     y = ts_thin$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2)

lines(x = ts_medium$frequency * 1e-3,
      y = ts_medium$TS,
      col = 'red',
      lty = 1,
      lwd = 2.5)

lines(x = ts_thick$frequency * 1e-3,
      y = ts_thick$TS,
      col = 'blue',
      lty = 1,
      lwd = 2.5)

legend("bottomright",
       c("0.5 mm shell", "1.0 mm shell", "2.0 mm shell"),
       lty = c(1, 1, 1),
       lwd = c(2.5, 2.5, 2.5),
       col = c('black', 'red', 'blue'),
       cex = 1.0)
```

## Effect of material properties

```{r fig.width = 8, fig.height = 4.5}
# Create spheres with different shell densities
sphere_light <- ess_generate(
  radius_shell = 10e-3, shell_thickness = 1e-3,
  sound_speed_shell = 3750, sound_speed_fluid = 1575,
  density_shell = 1500, density_fluid = 1077.3,  # Lighter shell
  K = 70e9, nu = 0.32
)

sphere_heavy <- ess_generate(
  radius_shell = 10e-3, shell_thickness = 1e-3,
  sound_speed_shell = 3750, sound_speed_fluid = 1575,
  density_shell = 4000, density_fluid = 1077.3,  # Heavier shell
  K = 70e9, nu = 0.32
)

# Calculate TS
sphere_light <- target_strength(sphere_light, frequency, "MSS_goodman_stern")
sphere_heavy <- target_strength(sphere_heavy, frequency, "MSS_goodman_stern")

# Extract results
ts_light <- extract(sphere_light, "model")$MSS_goodman_stern
ts_heavy <- extract(sphere_heavy, "model")$MSS_goodman_stern

# Plot comparison
par(oma = c(0, 0.25, 0, 0), mar=c(5, 6, 4, 2))
plot(x = ts_medium$frequency * 1e-3, 
     y = ts_medium$TS, 
     type = 'l',
     lty = 1, 
     lwd = 2.5,
     xlab = "Frequency (kHz)",
     ylab = expression(Target~strength~(dB~re.~1~m^2)),
     cex.lab = 1.3, 
     cex.axis = 1.2)

lines(x = ts_light$frequency * 1e-3,
      y = ts_light$TS,
      col = 'red',
      lty = 1,
      lwd = 2.5)

lines(x = ts_heavy$frequency * 1e-3,
      y = ts_heavy$TS,
      col = 'blue',
      lty = 1,
      lwd = 2.5)

legend("bottomright",
       c("Medium density (2565 kg/m³)", "Light shell (1500 kg/m³)", "Heavy shell (4000 kg/m³)"),
       lty = c(1, 1, 1),
       lwd = c(2.5, 2.5, 2.5),
       col = c('black', 'red', 'blue'),
       cex = 0.9)
```

# Model validity and applications

## Modal series vs. high-pass approximation

The modal series solution (`"MSS_goodman_stern"`) provides an exact solution valid
across all frequencies, while the high-pass approximation (`"high_pass_stanton"`)
is designed for high-frequency applications where the wavelength is much smaller
than the scatterer size.

## Biological applications

Elastic shelled spheres can represent various marine organisms:

- **Pteropods**: Pelagic gastropods with calcium carbonate and/or aragonite 
shells
- **Molluscs**: Benthic organisms with elastic shells
- **Coccolithophores**: Phytoplankton with calcite plates

The models implemented here provide the theoretical foundation for understanding
acoustic backscatter from these important marine organisms.

# Future development

Future enhancements may include:

- Additional shape options beyond spheres
- Temperature and pressure dependencies for material properties
- Frequency-dependent material properties
- Integration with empirical scattering databases
