---
title: "Introduction to the SDWBA acousticTS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using_SDWBA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## General Introduction

The purpose of `acousticTS` is the calculate the acoustic target strength (TS, dB re: m^2^) of a given scatterer, such as bubbles or euphausiids. Euphausiid TS is generally modelled using the distorted Born wave approximation (DWBA), which is a function of acoustic frequency (Hz), material properties (soundspeed, $h$, and density, $g$, contrast compared to ambient seawater), body morphometrics (shape, $r_{0}$, length, $L$, and curvature, $\rho_c$), and orientation relative to the sea surface ($\theta$). This model is used because it can be calculated for all frequencies, orientations, and animal shapes. The model treats each animal as a series of of bent cylinders that are generally weakly scattering ($\pm$ 5%). We use a special formulation of the model known as the SDWBA, which adds an additional phase deviation term ($\phi$) that accounts for imperfections in how sound scatters off of each cylinder. This model is mathematically expressed as:

$$
f_{bs}(\theta) = \sum_{j=1}^N [\frac{k_1~\rho_c}{4}~e^{(2~i~k_2~\rho_c)} \int M~e^{(-2~i~\vec{k_2}~cos~\beta_{tilt})} \frac{a_j J_1(2~k_2~a_j cos~ \beta_{tilt})}{cos~\beta_{tilt}}~d\beta_{tilt}] e^{(i~\phi_j)}
$$

where $N$ is the number of cylinders comprising each animal shape, $i$ is an imaginary number, $\vec{k_1}$ is the acoustic wave number of seawater adjusted for animal orientation $\theta$ whereby broadside incidence is 90$^\circ$/270$^\circ$, $\vec{k_2}$ is the acoustic wave number inside each cylinder adjusted for orientation, $\vec{r_0}$ is the position matrix which represents the x, y, and z coordinates of the animal shape, $a_j$ is the radius for cylinder $j$, $M$ is equal to $\frac{1}{gh^2} + \frac{1}{g-2}$, $g$ is the density contrast ($\frac{\rho_{animal}}{\rho_{seawater}}$), $h$ is the soundspeed contrast ($\frac{c_{animal}}{c_{seawater}}$), $J_1$ is the Bessel function of the first kind of order 1, $\rho_c$ is the radius of curvature, and $\phi_j$ is the stochastic phase variability term for each cylinder which provides a multiplicative term drawn from a standard normal distribution: $N(0, \phi_j)$. The TS of the scatterer is then calculated via $TS = 20~log_{10}(f_{bs}(\theta))$. 

## Package installation

The `acousticTS` package can be found on GitHub at https://github.com/brandynlucca/acousticTS and was built using R version 3.6.1. The package is actively updated, so updates and new features can be tracked via this repository. R is an open-source, statistical programming-language commonly used for data analysis across many fields. Updated versions of R can be installed on Windows, MacOS, and UNIX platforms from https://r-project.org. This package only has one dependency: `elliptic` (integration of complex numbers).

To install and call the latest version of `acousticTS` from the development GitHub repository: 

```{r message=F,warning=F,results="hide",eval=F}
#install.packages("devtools") #if needed, uncomment this line and install devtools
require(devtools) #call in devtools
install_github("brandynlucca/acousticTS")
```

Full documentation is provided for all functions (e.g., `?jl` provides the help documentation for the built-in spherical Bessel function of the first kind). 

## Model syntax and setup

The TS of an individual scatterer can be calculated using the `SDWBA(...)` (see: `?SDWBA`) function which requires a `fluid-filled scatterer (FFS)` (see: `?FFS`) class object, which is a S4 object that comprises the target's position matrix ($r_0$), radius ($a$), material properties ($g$, $h$), orientation ($\theta$), body curvature ($\rho_c$), body curve (a boolean `TRUE` or `FALSE`), body length ($L$), and number of cylinders ($N$). A sample shape can be called via `data(mcgehee)` (see: `?mcgehee`), which calls in the shape file and other model parameters reported in McGehee *et al.* (1998) for *Euphausia superba*, which has been commonly used for TS modelling efforts for krill. There are four ways to calculate TS using `SDWBA(...)`, which provides flexibility depending on users' needs, data format, and coding preferences:

1. Manually parameterizing the model.

```{r warnings=F,message=F}
#Let's call in the package
require(acousticTS)

#First we will create vectors for the x-, y-, and z-axis. 
x <- seq(1,10,1)*1e-3; y <- rep(0,10); z <- c(seq(1,5,1),rev(seq(1,5,1)))*1e-4

#Now we will assign a radius vector
a <- z/2

#Material properties
g <- 1.036; h <- 1.0279

#Fill out the remaining model parameters
c <- 1500 #sound speed, m/s
freq <- 120e3 #frequency, Hz

#Now let's calculate the TS of this hypothetical shape at 120 kHz and ambient sound speed of 1500 m/s
SDWBA(c=c, frequency=freq, x=x, y=y, z=z, a=a, g=g, h=h)

```


2. Manually creating a FFS-object using `FFSgenerate(...)` (see: `?FFSgenerate`) and `FFSwrite(...)` (see: `?FFSwrite`). 

```{r}
#We will use the same information we used before to generate our new scatterer
target <- FFSgenerate(x=x,y=y,z=z,a=a,g=g,h=h)

#We can look at the structure to get an idea about what the FFS scatterer object looks like
str(target)

#Now let's run the model on the object
SDWBA(target, c=c, frequency=freq)

#If we want to save this object for later modelling efforts, we can create a data.frame and write a *.csv file that is properly formatted
#Note that the FFSwrite(..., filename) defaults to saving this file in your current working directory. Name this file accordingly. 
#FFSwrite(target)
```

3. Calling in a *.csv formatted file to create a FFS-object using `FFSread(...)` (see: `?FFSread`).

```{r}
#We will call in a sample krill shape reported by McGehee et al. (1998) from the GitHub repository, but this can be applied to any *.csv file with the correct filepath. 
target_premade <- FFSread("https://raw.githubusercontent.com/brandynlucca/acousticTS/test-branch/data/mcgehee.csv")

#We can see that the file structure is the same as the previous "target" scatterer, just filled with different values
str(target_premade)

#Now let's run the model on the object
SDWBA(target_premade, c=c, frequency=freq)
```

4. Using both arbitrary and pre-generated FFS-object parameters. 

```{r}
#Let's say we are interested in a very different set of material properties!
SDWBA(target_premade, c=c, frequency=freq, g=1.01, h=1.005)
```

## FFS-object manipulation

In the previous section it was assumed that all of the scatterers we not curved and were straight from rostrum-to-telson. In these cases, a more traditional form of the model was used that discounts $\rho_c$. An animal can be curved by either setting `SDWBA(..., curve=T)` or using the `Shapely(...)` (see: `?Shapely(...)`). If `curve=TRUE`, the default $\rho_c$ (i.e., `pc`) is 3.0 unless it is otherwise defined. 

```{r}
#Let's take a look at the str of our target. We can see that "curve" is set to False, and the radius of curvature (pc) is set to 0.0. 
str(target_premade)

#So we can force the object to do our bidding and curve it.
target_curved <- Shapely(target_premade, curve=T, pc=3.3)

#Now let's calculate the TS using both of the described methods, which will yield the same answer
SDWBA(target_curved, c=c, frequency=freq)
SDWBA(target_premade, c=c, frequency=freq, curve=T, pc=3.3)
```

The `Shapely(...)` function can also be used to change the orientation angle (`Shapely(..., theta=angle)`) and resize both the $r_0$ matrix and $a$ vector based on a new length (`Shapely(..., length=new_length)`). The identical alternative is available by manually inputting a new orientation value into `SDWBA(...)`, but the shape manipulation requires `Shapely(...)`. For the purpose of convenience, the `degrad(...)` (see: `?degrad`) function is available for readily converting degrees to radians, and vice versa. 

```{r}
#First let's change the orientation of the object
target_angle <- Shapely(target_curved, theta=pi)
SDWBA(target_angle, c=c, frequency=freq)
SDWBA(target_curved, c=c, frequency=freq, theta=pi)

#If our orientation values are in degrees, we can use degrad(...) to convert! 
angle_radians <- degrad(90,"deg") #degrees to radians
angle_radians
angle_degrees <- degrad(angle_radians,"rad") #degrees to radians
angle_degrees

#We can also resize our target shape; the new length is in m
target_resize <- Shapely(target_angle, length=30e-3)
target_resize

#So we have resized our target from a 41 mm krill to 30 mm, which adjusted both the position matrix, rpos, and our radius vector, a.
SDWBA(target_resize, c=c, frequency=freq)

```


## Stochastic/vectorized model inputs

Up until now, we have not added in the phase deviation term ($\phi$) that is set via `SDWBA(..., phase=0.0)`. As noted, this sets the standard deviation term to draw from a standard normal distribution that is drawn from `rnorm(1,0,phase)`. We can see that this can have a negligible-to-large effect depending on how large $\phi$ is set. 

```{r}
set.seed(1000)
#Normal DWBA, phase = 0.0
SDWBA(target_resize, c=c, frequency=freq)
#phase = 0.04
SDWBA(target_resize, c=c, frequency=freq, phase=0.04)
#phase = 0.07
SDWBA(target_resize, c=c, frequency=freq, phase=0.07)
#phase = 0.22
SDWBA(target_resize, c=c, frequency=freq, phase=0.22)
#phase = 0.50
SDWBA(target_resize, c=c, frequency=freq, phase=0.50)
#phase = 0.77
SDWBA(target_resize, c=c, frequency=freq, phase=0.77)
```

Because this is a stochastic variable, it makes more sense to simulate over more than a single iteration. This can be achieved using `SDWBA.sim(...)` (see: `?SDWBA.sim`), specifically defining `SDWBA.sim(..., nrep=NULL)`:

```{r}
#Let's generate 10 iterations using the phase deviation, phi, of 0.77. This will print a bunch of values with their respective parameters.
set.seed(1000)
SDWBA.sim(target_resize, c=c, frequency=freq, phase=0.77, nrep=5)

#This can also be parallelized, if desired and available to your computer by setting SDWBA.sim(..., parallel=T) and defining the number of cores you'd like to use via SDWBA.sim(..., n.cores=number); the default number of cores will revert to the number of cores available to your computer.
set.seed(1000)
SDWBA.sim(target_resize, c=c, frequency=freq, phase=0.77, nrep=5, parallel=T, n.cores=6)
```

Note that `SDWBA.sim(...)` requires a FFS-class object input. Of course not all targets/animals comprise the same model parameters, or comprise a distribution of values such as various orientations, body curvatures, and material properties. We can also use `SDWBA.sim(...)` to draw from distributions for these other parameters:

```{r warning=F}
#Using our same curved shape above, let's play around with a few distributions
n <- 5 #number of animal parameters we want to simulate
lengths <- rnorm(n,mean=30,sd=5)*1e-3 #Normal distribution of lengths
g <- rnorm(n,mean=target_resize@g,sd=1e-3) #Normal distribution for g
h <- rnorm(n,mean=target_resize@h,sd=1e-3) #Normal distribution for h
theta <- c(pi/2,pi/4) #Play around with two different orienation angles
pd <- 0.77 #phase deviation

#Now let's simulate over every combination of these values
ts_sim <- SDWBA.sim(target_resize, c=c, frequency=freq, phase=pd, g=g, h=h, length=lengths, theta=theta, parallel=T)
head(ts_sim)

#We can then look at the distribution of values to see our simulated TS results! 
require(ggplot2)

ggplot(data=ts_sim, aes(x=TS)) + geom_histogram(fill="red", color="black") +
  theme_bw() + theme(text=element_text(size=16), axis.text=element_text(size=14, color="black"),
                     panel.grid=element_blank()) +
  labs(x=expression(paste("Target strength (dB re: 1 ",m^2,")")),
       y="Frequency")
```

## Example: TS-Frequency spectrum comparison between curved and non-curved animal

```{r}
#We will compare the TS-frequency spectra of the same target when it is curved and non-curved
curve = c(T,F) #curved and non-curved
freq <- seq(10,300,1)*1e3 #frequencies to simulate, Hz
theta <- pi/2 #broadside incidence
ts_freq <- SDWBA.sim(target_resize, c=c, frequency=freq, curve=curve, theta=theta, parallel=T)

#And we can plot!
ggplot(data=ts_freq, aes(y=TS,x=frequency)) + geom_path(size=2) +
  facet_wrap(~curve)
                                                        
                                                        
  theme_bw() + theme(text=element_text(size=16), axis.text=element_text(size=14, color="black"),
                     panel.grid=element_blank()) +
  scale_color_manual(values=c("red","blue")) +
  labs(x=expression(paste("Target strength (dB re: 1 ",m^2,")")),
       y="Frequency")
```
