---
title: "KRM implementation"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{KRM implementation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  status = c("benchmarked", "validated"),
  pages = c(
    Overview = "index.html",
    Implementation = "krm-implementation.html",
    Theory = "krm-theory.html"
  )
)
```

The `acousticTS` package uses object-based scatterers, so the KRM workflow follows the same broad structure used elsewhere in the package: define the geometry, attach material meaning to the body and any internal inclusion, evaluate the model over the desired frequency range, and then inspect the returned response carefully enough to understand what part of the target is driving it. For KRM, the main structural choice is whether the target is represented as a fluid-like body only or as a body plus an internal strongly contrasting inclusion such as a swimbladder.

This matters because KRM is not just a body-shape model. It is a composite body-plus-inclusion model. Readers should therefore treat object construction as part of the scientific interpretation rather than just a data-formatting step.

## Fish-like object generation

The example below creates a simple body shape and an embedded swimbladder shape with `arbitrary()` and passes both to `sbf_generate()`.

```{r}
library(acousticTS)

body_shape <- arbitrary(
  x_body = c(0.00, 0.04, 0.09, 0.15, 0.22, 0.28),
  w_body = c(0.00, 0.018, 0.028, 0.030, 0.018, 0.00),
  zU_body = c(0.000, 0.010, 0.016, 0.017, 0.010, 0.000),
  zL_body = c(0.000, -0.010, -0.016, -0.017, -0.010, 0.000)
)

bladder_shape <- arbitrary(
  x_bladder = c(0.06, 0.10, 0.14, 0.18, 0.21),
  w_bladder = c(0.00, 0.008, 0.012, 0.008, 0.00),
  zU_bladder = c(0.000, 0.004, 0.006, 0.004, 0.000),
  zL_bladder = c(0.000, -0.004, -0.006, -0.004, 0.000)
)

fish_object <- sbf_generate(
  body_shape = body_shape,
  bladder_shape = bladder_shape,
  density_body = 1070,
  sound_speed_body = 1570,
  density_bladder = 1.2,
  sound_speed_bladder = 340,
  theta_body = pi / 2,
  theta_bladder = pi / 2
)

fish_object
```

This object already encodes the key KRM assumptions. The body and bladder are described separately, their material properties are assigned separately, and the orientation of each is carried explicitly. That separation is important because the KRM assigns different scattering logic to weakly contrasting body tissue and to the strongly contrasting internal inclusion. If the geometry or placement of the bladder is unrealistic, the model may still run, but the interpretation of the result will no longer match the intended biological target.

## Calculating a TS-frequency spectrum

```{r}
frequency <- seq(18e3, 120e3, by = 6e3)

fish_object <- target_strength(
  object = fish_object,
  frequency = frequency,
  model = "krm"
)
```

This call evaluates the composite KRM response over the requested frequency grid. As with other approximations, it is worth choosing the grid with the interpretation in mind. A broader sweep is useful for seeing how the body-only and bladder-dominated regimes compare, while a narrower and finer grid may be more useful when the question is how strongly the bladder alters the spectrum over a particular band of interest.

## High- and low-ka swimbladder assumptions

For fluid-only body targets, `krm_variant` is ignored because only the body Kirchhoff term is evaluated. The argument matters only for combined body-plus-swimbladder targets, where the model must decide what surrounding medium is used in the swimbladder terms (Clay and Horne 1994; Horne and Jech 1999).

In the high-`ka` swimbladder Kirchhoff term, `acousticTS` uses the empirical factors

$$
  A_{SB,s} = \frac{ka_s}{ka_s + 0.083},
  \qquad
  \psi_{p,s} = \frac{ka_s}{40 + ka_s} - 1.05,
$$

and the local oscillatory term

$$
  \mathcal{L}_{SB,s} \propto
  A_{SB,s}
  \big[(k_H a_s + 1)\sin\theta\big]^{1/2}
  e^{-i(2k_H v_s + \psi_{p,s})}\Delta u_s,
$$

where \(k_H\) is the wavenumber assumed for the high-`ka` surrounding medium. In the low-`ka` breathing-mode term, the relevant acoustic size is

$$
  ka_e = k_L a_e,
$$

where \(k_L\) is the wavenumber assumed for the low-`ka` surrounding medium.

The three named variants therefore encode the modelling assumption directly:

1. `lowcontrast`: apply the low-contrast approximation \(k_B \approx k\) in both swimbladder regimes, so \(k_H = k_L = k\).
2. `mixed`: use the body-medium wavenumber for the high-`ka` swimbladder term but the low-contrast approximation for the low-`ka` breathing mode, so \(k_H = k_B\) and \(k_L = k\).
3. `body_embedded`: use the body-medium wavenumber in both swimbladder regimes, so \(k_H = k_L = k_B\). This is the most literal body-embedded reading of Clay and Horne (1994).

```{r fig.width = 6, fig.height = 4.5}
data(sardine)

sardine_freq <- seq(12e3, 400e3, by = 2e3)

sardine_lowcontrast <- target_strength(
  object = sardine,
  frequency = sardine_freq,
  model = "krm",
  density_sw = 1030,
  sound_speed_sw = 1490,
  krm_variant = "lowcontrast"
)

sardine_mixed <- target_strength(
  object = sardine,
  frequency = sardine_freq,
  model = "krm",
  density_sw = 1030,
  sound_speed_sw = 1490,
  krm_variant = "mixed"
)

sardine_body_embedded <- target_strength(
  object = sardine,
  frequency = sardine_freq,
  model = "krm",
  density_sw = 1030,
  sound_speed_sw = 1490,
  krm_variant = "body_embedded"
)

sardine_low_df <- extract(sardine_lowcontrast, "model")$KRM
sardine_mixed_df <- extract(sardine_mixed, "model")$KRM
sardine_body_df <- extract(sardine_body_embedded, "model")$KRM

par(oma = c(0, 0.25, 0, 0), mar = c(5, 6, 4, 2))
plot(
  sardine_low_df$frequency * 1e-3,
  sardine_low_df$TS,
  type = "l",
  lwd = 2.5,
  col = "black",
  xlab = "Frequency (kHz)",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  cex.lab = 1.2,
  cex.axis = 1.1
)
lines(
  sardine_body_df$frequency * 1e-3,
  sardine_body_df$TS,
  col = "firebrick3",
  lwd = 2.5
)
lines(
  sardine_mixed_df$frequency * 1e-3,
  sardine_mixed_df$TS,
  col = "dodgerblue3",
  lwd = 2,
  lty = 2
)
legend(
  "bottomleft",
  legend = c("lowcontrast", "body_embedded", "mixed"),
  col = c("black", "firebrick3", "dodgerblue3"),
  lwd = c(2.5, 2.5, 2),
  lty = c(1, 1, 2),
  bty = "n"
)
```

For the bundled `sardine` target over `12-400 kHz`, the `mixed` and `body_embedded` curves are numerically identical, so the practical split is between the default `lowcontrast` spectrum and the body-medium alternative.

## Extracting model results

Model results can be extracted either visually or directly through `extract()`.

### Plotting results

```{r fig.width = 6}
plot(fish_object, type = "shape")
plot(fish_object, type = "model")
```

### Accessing results

```{r}
krm_results <- extract(fish_object, "model")$KRM
head(krm_results)
```

At this stage, it is worth confirming more than the presence of a plotted curve. Readers should check that the returned frequencies match the request, that the target-strength scale is plausible for the assumed fish size and material contrasts, and that the result is being interpreted as a coherent composite response rather than as an isolated body or isolated bladder term.

## Comparing body-only and body-plus-swimbladder cases

The most common practical comparison in KRM work is the contribution of the swimbladder relative to the fluid-like body.

```{r fig.width = 8, fig.height = 4.5}
body_only_object <- fls_generate(
  shape = body_shape,
  density_body = 1070,
  sound_speed_body = 1570,
  theta_body = pi / 2
)

body_only_object <- target_strength(
  object = body_only_object,
  frequency = frequency,
  model = "krm"
)

ts_body_only <- extract(body_only_object, "model")$KRM
ts_body_and_bladder <- extract(fish_object, "model")$KRM

par(oma = c(0, 0.25, 0, 0), mar = c(5, 6, 4, 2))
plot(
  x = ts_body_only$frequency * 1e-3,
  y = ts_body_only$TS,
  type = "l",
  lty = 1,
  lwd = 2.5,
  xlab = "Frequency (kHz)",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  cex.lab = 1.3,
  cex.axis = 1.2,
  ylim = c(-70, -20)
)
lines(
  x = ts_body_and_bladder$frequency * 1e-3,
  y = ts_body_and_bladder$TS,
  col = "firebrick3",
  lty = 1,
  lwd = 2.5
)
legend(
  "bottomright",
  legend = c("Body only", "Body plus swimbladder"),
  col = c("black", "firebrick3"),
  lwd = 2.5,
  lty = c(1, 1),
  cex = 1.0,
  bty = "n"
)
```

This comparison is often the most informative first diagnostic because it isolates the physical reason the KRM exists. Holding the body geometry fixed while adding or removing the internal inclusion shows whether the total response is dominated by weakly scattering tissue, by the strong internal contrast, or by coherent interaction between both contributions.

For practical KRM work, the first inputs to revisit are usually:

1. the body and bladder geometry, because the model assumes a sensible locally cylindrical segmentation,
2. the relative placement and scale of the bladder inside the body, because that strongly affects the composite interpretation, and
3. the material contrasts assigned to both regions, because they control whether the body remains weakly scattering and whether the bladder remains the dominant internal inclusion.

For more realistic biological examples, the bundled `cod` and `sardine` datasets are the natural next objects to inspect because they already package body and swimbladder geometry in the format expected by the model.

## Modal-series benchmark comparisons

The primary KRM implementation benchmark for the canonical gas-filled and weakly scattering shapes is the published modal-series reference solution for each isolated canonical case. The summary table below keeps that benchmark visible without mixing it with the separate software-compatibility checks for body-plus-swimbladder targets.

| Canonical case | Max abs. delta TS (dB) | Mean abs. delta TS (dB) | Elapsed (s) |
|:--|--:|--:|--:|
| Gas sphere | 7.91469 | 6.04116 | 0.01 |
| Weakly scattering sphere | 11.21090 | 0.47451 | 0.15 |
| Gas prolate spheroid | 7.36054 | 4.62751 | 0.00 |
| Weakly scattering prolate spheroid | 14.27193 | 0.46239 | 0.00 |
| Gas cylinder | 7.35632 | 6.19094 | 0.02 |
| Weakly scattering cylinder | 23.22115 | 0.59807 | 0.02 |

The modal-series table above remains the implementation benchmark for the canonical KRM targets.

## Bundled fish compatibility checks

The bundled fish objects serve a different benchmark role from the canonical modal-series targets above. There is no exact modal-series reference for these segmented body-plus-swimbladder geometries, so they are used only to check software-to-software agreement for the `krm_variant` branches on realistic fish shapes.

| Object | Branch or comparison | Mean abs. delta TS (dB) | Max abs. delta TS (dB) |
|:--|:--|--:|--:|
| `sardine` | `acousticTS lowcontrast` vs `KRMr` | 1.58e-14 | 1.42e-13 |
| `sardine` | `acousticTS lowcontrast` vs `echoSMs` low-contrast spectrum | 3.31e-14 | 5.54e-13 |
| `sardine` | `acousticTS lowcontrast` vs NOAA applet spectrum | 0.00262 | 0.00496 |
| `sardine` | `KRMr` vs `echoSMs` low-contrast spectrum | 3.85e-14 | 5.54e-13 |
| `sardine` | `KRMr` vs NOAA applet spectrum | 0.00262 | 0.00496 |
| `sardine` | `echoSMs` low-contrast spectrum vs NOAA applet spectrum | 0.00262 | 0.00496 |
| `sardine` | `acousticTS body_embedded` vs `echoSMs` body-embedded spectrum | 3.37e-14 | 1.85e-13 |
| `sardine` | `acousticTS mixed` vs `echoSMs` body-embedded spectrum | 3.37e-14 | 1.85e-13 |
| `sardine` | `acousticTS mixed` vs `acousticTS body_embedded` | 0.00000 | 0.00000 |
| `cod` | `acousticTS lowcontrast` vs `KRMr` | 1.33e-14 | 7.82e-14 |
| `cod` | `acousticTS lowcontrast` vs `echoSMs` low-contrast spectrum | 5.36e-15 | 3.55e-14 |
| `cod` | `acousticTS lowcontrast` vs NOAA applet spectrum | 0.00262 | 0.00498 |
| `cod` | `KRMr` vs `echoSMs` low-contrast spectrum | 1.46e-14 | 8.53e-14 |
| `cod` | `KRMr` vs NOAA applet spectrum | 0.00262 | 0.00498 |
| `cod` | `echoSMs` low-contrast spectrum vs NOAA applet spectrum | 0.00262 | 0.00498 |
| `cod` | `acousticTS body_embedded` vs `echoSMs` body-embedded spectrum | 6.16e-15 | 5.68e-14 |
| `cod` | `acousticTS mixed` vs `echoSMs` body-embedded spectrum | 6.16e-15 | 5.68e-14 |
| `cod` | `acousticTS mixed` vs `acousticTS body_embedded` | 0.00000 | 0.00000 |

Those values make the branch behavior explicit:

1. `lowcontrast` remains the default because it reproduces the low-contrast software family on both bundled fish objects.
2. `body_embedded` remains the literal body-medium alternative and lines up with the corresponding body-embedded spectrum.
3. `mixed` is still exposed because it is a distinct literature convention, even though on both bundled fish frequency grids it collapses numerically onto `body_embedded`.

So the two benchmark roles are now explicit:

1. canonical modal-series tables for isolated-shape KRM implementation checks, and
2. bundled-fish compatibility checks for the body-plus-swimbladder convention choice.


