---
title: "Validation and Benchmark Reproduction"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Validation and Benchmark Reproduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# What this page establishes

Validation is evidence that a model reproduces an independent result within a
stated scope. It is not established by a successful function call or a
plausible-looking curve. `acousticTS` uses stored canonical benchmarks,
comparisons with independent implementations, numerical checks, and regression
tests. These forms of evidence answer different questions and should not be
treated as interchangeable [@Jech_2015; @echoSMs_software].

![Validation proceeds from object and unit checks through known limits and published or independent comparisons.](validation-ladder.svg)

The ladder is ordered deliberately. A benchmark comparison is not interpretable
until geometry, units, boundary condition, medium properties, orientation, and
output normalization have been matched.

# Four distinct checks

## Setup checks

Confirm that the object represents the intended geometry and physical target.
Inspect dimensions, units, component placement, material properties, medium
properties, boundary condition, frequency, and orientation. This catches many
large discrepancies before they are attributed to a solver.

## Numerical verification

Check that the implemented equations are being solved consistently. Depending
on the model, this can include convergence with modal truncation, quadrature
order, geometric resolution, precision, or iterative tolerance. Known limiting
behavior and agreement between mathematically equivalent formulations also
belong here.

## Benchmark or external validation

Reproduce a case whose target definition and expected result are documented.
Canonical data from inter-model studies are useful for this purpose
[@Jech_2015]. Independent software comparisons provide stronger separation from
the package implementation when the two codes do not share the same numerical
path [@echoSMs_software; @ZooScatR_software; @SphereTS_software]. Agreement
applies only to the geometry, boundary, parameter range, and output quantity
that were compared.

## Regression testing

A regression test detects whether package output changed. It does not by itself
show that the stored output is physically correct. Regression values are most
useful after the underlying case has been tied to theory, a benchmark, or an
independent calculation.

# Package validation status

The family badges summarize the documented evidence currently registered in the
package:

```{r validation_policy, echo=FALSE, results='asis'}
acousticTS:::.validation_status_policy()
```

```{r validation_status_table, echo=FALSE}
knitr::kable(
  acousticTS:::.validation_family_status_table(),
  align = c("l", "l", "l")
)
```

These labels are scope statements. A family can be benchmarked and experimental
at the same time, or validated for some branches while other public branches
remain outside the external comparison.

<details>
<summary>Show the evidence recorded for each family</summary>

```{r validation_evidence_table, echo=FALSE}
knitr::kable(
  acousticTS:::.validation_evidence_table(),
  align = c("l", "l", "l", "l", "l")
)
```

</details>

Use the relevant model's implementation page for the actual comparison setup,
numerical controls, plots, and limitations. The registry is an index of that
evidence, not a substitute for it.

# The bundled benchmark data

[`benchmark_ts`](../../reference/benchmark_ts.html) contains target-strength
spectra for canonical sphere, finite-cylinder, and prolate-spheroid cases, plus
angle spectra for the cylinder and prolate spheroid. The data follow the target
definitions assembled for the comparison by @Jech_2015 and the associated
`echoSMs` resources [@echoSMs_software].

```{r}
library(acousticTS)

data(benchmark_ts)

names(benchmark_ts)
names(benchmark_ts$frequency_spectra)
names(benchmark_ts$frequency_spectra$sphere)
```

The benchmark object stores expected outputs, not a complete self-executing
test. A reproduction still has to rebuild the documented geometry and physical
conditions explicitly.

# Reproduce a canonical sphere case

The following example rebuilds the 10 mm-radius rigid-sphere frequency
benchmark using only exported package functions. It is not evaluated while the
site is built because the purpose here is to show the reproducible check rather
than add another full model run to documentation rendering:

```{r eval=FALSE}
data(benchmark_ts)

frequency <- benchmark_ts$frequency_spectra$index$frequency
reference <- benchmark_ts$frequency_spectra$sphere$fixed_rigid

sphere_obj <- gas_generate(
  shape = sphere(radius_body = 0.01, n_segments = 60)
)

sphere_obj <- target_strength(
  object = sphere_obj,
  frequency = frequency,
  model = "sphms",
  boundary = "fixed_rigid",
  density_sw = 1026.8,
  sound_speed_sw = 1477.3
)

predicted <- extract(sphere_obj, "model")$SPHMS$TS
delta_dB <- predicted - reference

c(
  max_abs_dB = max(abs(delta_dB)),
  rmse_dB = sqrt(mean(delta_dB^2))
)

all.equal(predicted, reference, tolerance = 1e-2)
```

The tolerance is part of the test definition. It should reflect the benchmark
precision, interpolation, numerical method, and scientific purpose. It should
not be adjusted only until a failing comparison passes.

The [SPHMS](../sphms/sphms-implementation.html),
[FCMS](../fcms/fcms-implementation.html), and
[PSMS](../psms/psms-implementation.html) implementation pages provide their
model-specific benchmark details. Some PSMS comparisons require particular
precision and integration settings, which is why benchmark reproduction must
record numerical controls as well as physical inputs.

# Choose the comparison quantity

Compare the quantity that the reference result actually reports:

- Use `TS` for agreement in reported decibels.
- Use `sigma_bs` when relative or absolute scattering strength in the linear
  domain is the question.
- Use complex amplitude when phase, interference, or coherent combination is
  being tested.

Errors near a deep null can appear large in decibels even when the absolute
linear-domain discrepancy is small. Conversely, agreement in `TS` does not test
phase. Report the comparison domain with the metric and tolerance.

# Check numerical stability around the benchmark

Matching one stored curve is not enough when the result is sensitive to a
numerical control. Repeat the calculation while varying the controls that apply
to the model:

- shape or boundary discretization,
- modal truncation,
- quadrature order,
- floating-point precision,
- iterative tolerance, and
- frequency or angle resolution used to compare curves.

Change one control at a time. A result is numerically stable when further
refinement changes the quantity of interest by less than a stated tolerance.
Do not use agreement at one hand-picked setting to conceal instability nearby.
See [Numerical Methods and Special
Functions](../numerical-foundations/numerical-foundations.html) for the package's
underlying numerical components.

# Record enough information to reproduce the check

A useful validation record includes:

1. the package version or commit,
2. the benchmark source and citation,
3. target geometry and units,
4. material, medium, and boundary definitions,
5. frequency and angle grids,
6. model and numerical options,
7. the output domain, error metric, and tolerance, and
8. the code and input hashes used to generate any stored comparison figure or
   table.

That record distinguishes a reproducible benchmark from a plot that merely
resembles a published result. When a model is used outside the registered
comparison scope, describe the result as an extrapolation from validated or
benchmarked behavior rather than extending the validation claim implicitly.
