---
title: "Validation and benchmark reproduction"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Validation and benchmark reproduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

The validation workflow on this page is grounded in inter-model benchmark comparisons and open-software reproductions [@jech_etal_2015; @demer_validation_2003; @gastauer_zooscatrspan_2019; @sven_gastauer_svengastauerkrmr_2025; @betcke_bempp-cl_2021].



Modeling packages become much more useful when users can distinguish a physically surprising result from a setup error. Validation pages help provide that context. In acousticTS, validation is not a single check with a single threshold. It is a layered workflow that begins with object and unit sanity, then moves to canonical benchmark reproduction, and finally asks whether the numerical behavior is stable enough that any remaining discrepancy deserves physical interpretation.

![Validation ladder](validation-ladder.svg)

That layered view matters because benchmark comparisons are only informative when the target definition, acoustic assumptions, numerical settings, and reporting scale are already aligned. A mismatch against a stored benchmark curve does not automatically diagnose a broken model. It may instead reflect a different boundary condition, a different frequency grid, a different medium, a different orientation convention, or a comparison being carried out in a quantity that is not the one the benchmark was intended to validate.

# What the benchmark resources are for

The package includes `benchmark_ts`, which provides stored benchmark curves for several canonical model-and-geometry combinations. Together with the bundled example scatterers and the theory vignettes, it provides a documented basis for regression-style checks. That makes the dataset more than a teaching convenience. It is also part of the package's broader numerical validation strategy.

```{r}
library(acousticTS)

data(benchmark_ts)

names(benchmark_ts$frequency_spectra)
head(benchmark_ts$frequency_spectra$index$frequency)
```

The benchmark cases currently emphasize canonical exact or near-exact comparison problems: `SPHMS` sphere cases with several boundary conditions, `FCMS` finite-cylinder cases, and `PSMS` prolate-spheroid cases. That combination is useful because it spans several canonical geometries and several different numerical mechanisms. In practice, these resources let a user ask two related but different questions. The first is whether a known reference problem can be reproduced. The second is whether a new workflow still behaves consistently when anchored to a trusted canonical case.

# Validation registry snapshot

The package now also keeps a small internal validation registry so the family-page badges, model-library summaries, and validation evidence can be drawn from one source rather than maintained separately.

## Family statuses

```{r validation_status_table, echo=FALSE}
knitr::kable(
  acousticTS:::.validation_family_status_table(),
  align = c("l", "l", "l")
)
```

## Benchmark evidence

```{r benchmark_evidence_table, echo=FALSE}
knitr::kable(
  acousticTS:::.validation_evidence_table(type = "benchmark"),
  align = c("l", "l", "l", "l", "l")
)
```

## External-comparison evidence

```{r external_evidence_table, echo=FALSE}
knitr::kable(
  acousticTS:::.validation_evidence_table(type = "external"),
  align = c("l", "l", "l", "l", "l")
)
```

# What validation means in this package

Validation in acousticTS is best understood as at least four related layers of evidence.

First, the object and setup have to be internally consistent. The geometry must be the one the user thinks was created, the material properties must be in the correct units, the model must be appropriate for the target class and boundary interpretation, and the frequency grid and orientation convention must match the intended analysis. If any of that is wrong, benchmark disagreement says very little.

Second, the workflow should reproduce known theoretical or canonical limits whenever those are available. For some models that means checking against an exact modal-series benchmark. For others it means checking whether the limiting behavior is consistent with the approximation regime documented in the theory page.

Third, the numerical solution has to be stable enough that the reported agreement is meaningful. Some comparisons are only trustworthy after truncation settings, precision choices, quadrature controls, or object resolution have been perturbed and found not to change the result materially.

Fourth, the reporting quantity has to match the comparison question. A benchmark assessed in `TS` is not necessarily saying the same thing as a benchmark assessed in `sigma_bs`, and neither is equivalent to direct comparison of the complex scattering amplitude.

# Layer 1: geometry and setup checks

The first layer of validation is simple but essential. Confirm the shape is what you think it is, confirm that the material properties and units are sensible, confirm that the selected model is compatible with the target class and boundary interpretation, and confirm that the frequency grid and orientation convention match the intended benchmark. This layer is often enough to explain large discrepancies. Many apparent model failures are actually mismatches in units, contrasts, orientation, or boundary-condition naming.

This is also the stage where readers should decide whether a benchmark is even comparable to the problem they are running. A rigid-sphere benchmark is not a useful reference for a fluid-filled sphere setup, and a benchmark built in seawater is not a direct check on a target rebuilt with different ambient properties unless the comparison has been intentionally nondimensionalized.

# Layer 2: benchmark reproduction

The second layer is to reproduce a known benchmark curve or stored regression case. The bundled resources and the package tests provide the clearest examples of this workflow. For spheres, one can build the canonical sphere case, run `SPHMS`, and compare the resulting `TS` vector against the stored benchmark curve.

```{r eval = FALSE}
data(benchmark_ts)

frequency <- benchmark_ts$frequency_spectra$index$frequency
density_sw <- 1026.8
sound_speed_sw <- 1477.3

scatterer_ess <- fixture_sphere("fixed_rigid")

scatterer_ess <- target_strength(
  scatterer_ess,
  frequency = frequency,
  model = "sphms",
  boundary = "fixed_rigid",
  density_sw = density_sw,
  sound_speed_sw = sound_speed_sw
)

all.equal(
  extract(scatterer_ess, "model")$SPHMS$TS,
  benchmark_ts$frequency_spectra$sphere$fixed_rigid,
  tolerance = 1e-2
)
```

The cylinder and prolate-spheroid benchmark workflows follow the same broad pattern with `FCMS` and `PSMS`, respectively. For the prolate-spheroid case, the benchmark setup also shows something important: faithful reproduction can require model-specific numerical settings such as `precision = "quad"` or `simplify_Amn = FALSE`. In other words, a model may be mathematically correct and still miss a benchmark if the numerical configuration is not the one the comparison requires.

# Layer 3: choosing the comparison quantity

Benchmark agreement should be interpreted in the quantity that answers the validation question. When the goal is to reproduce the exact reported target-strength curve from a publication or stored package resource, comparison in `TS` is often appropriate because the benchmark itself is already expressed in decibels. In that setting, absolute dB differences are directly interpretable and are often the most convenient regression metric.

When the goal is to assess scattering strength as a physical quantity, or when the benchmark is really about backscattering magnitude rather than reported target strength, comparison in the linear domain is usually more informative. That means working with `sigma_bs` or, when phase matters, with `f_bs`. Two solutions that differ modestly in `TS` can represent a much larger relative disagreement in `sigma_bs` at low-scattering portions of the curve. The converse can also happen: a visually noticeable dB mismatch near a strong peak may correspond to a smaller relative linear-domain error than a reader expects.

For that reason, validation metrics such as MAD, RMSE, or maximum absolute error should be interpreted together with the domain in which they were computed. A dB-domain RMSE rewards agreement in the reported target-strength scale. A linear-domain RMSE rewards agreement in physical scattering magnitude. Neither dominates the other universally. The right choice depends on whether the benchmark is being used as a reporting check, a physical-strength check, or a numerical-regression check.

# Layer 4: numerical stability checks

Benchmark reproduction is necessary, but it is not the whole story. Some workflows also need local numerical checks such as increasing object resolution, changing truncation or integration controls, checking sensitivity to precision settings, and confirming that oscillatory structure is physically consistent rather than an artifact. This is especially relevant for models with heavier numerical machinery, such as `PSMS` or boundary-value problems involving multiple coupled coefficients.

A benchmark should therefore be thought of as an anchor, not as a substitute for local numerical inspection. If a curve agrees with a benchmark only at one specific precision or one specific truncation setting but changes materially under modest perturbation, the agreement should be treated cautiously. Conversely, a small residual mismatch may be acceptable when the solution is numerically stable and the comparison is known to involve different interpolation, reporting, or precision conventions.

# How to use the bundled benchmarks well

The bundled benchmark dataset is most useful when it is treated as a regression reference rather than just a plotting convenience. Good uses include checking a local modification to a model implementation, confirming that a custom object-construction workflow reproduces a known canonical case, verifying that changes in numerical options do not materially alter a trusted solution, and teaching the difference between a setup error and a model limitation.

Less useful uses include comparing a benchmark curve to a target built with different physical assumptions and then interpreting the mismatch as a numerical problem. A benchmark is only informative when the target definition and the acoustic assumptions are genuinely aligned. Before treating disagreement as evidence of model failure, it is worth confirming that the benchmark and the reproduced case actually describe the same geometry, medium, boundary condition, orientation convention, and reported quantity.

# A practical validation ladder

For routine work, a conservative sequence is usually best.

1. Plot the geometry and confirm that the object matches the intended canonical case.
2. Inspect the material properties, units, boundary interpretation, and medium parameters.
3. Run one deterministic model call and inspect the extracted outputs in the quantity the benchmark actually reports.
4. Compare against a canonical or benchmark case when one is available.
5. Vary one numerical control at a time if the result still looks suspicious.
6. Only after those checks, decide whether any remaining disagreement is physical, numerical, or simply a difference in reporting scale.

That ladder avoids blaming the mathematics before checking the setup, and it avoids over-interpreting benchmark mismatches that are really mismatches of assumptions.

# Related reading

- [Working with real example data](../example-data/example-data.html)
- [Numerical foundations and special functions](../numerical-foundations/numerical-foundations.html)
- [FAQ and troubleshooting](../faq-troubleshooting/faq-troubleshooting.html)
