---
title: "Bent cylinder modal series solution"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Bent cylinder modal series solution}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  status = c("experimental", "unvalidated"),
  pages = c(
    Overview = "index.html",
    Implementation = "bcms-implementation.html",
    Theory = "bcms-theory.html"
  )
)
```

The bent-cylinder modal-series solution is available through `target_strength(..., model = "bcms")`. For the straight branch, the implementation should reduce exactly to the finite-cylinder modal-series solution. For the bent branch, it should reduce to the straight finite-cylinder result multiplied by the Fresnel coherence factor implied by Stanton (1989, Eq. 25-26).

That is the implementation check used here. No separate public local BCMS package was identified, so the comparison is carried out against the governing straight-cylinder modal series and the direct bent-cylinder coherence construction.

# Reference case

The comparison uses a weakly scattering liquid-filled cylinder with:

- length `10.5 mm`
- radius `1 mm`
- curvature ratio `rho_c / L = 1.5` for the bent case
- density contrast `g = 1.0357`
- sound-speed contrast `h = 1.0279`
- broadside incidence
- `12-400 kHz` in `2 kHz` steps

The straight and bent objects are:

```{r}
library(acousticTS)

density_sw <- 1026.8
sound_speed_sw <- 1477.3

straight_object <- fls_generate(
  shape = cylinder(
    length_body = 10.5e-3,
    radius_body = 1e-3,
    n_segments = 401
  ),
  density_body = density_sw * 1.0357,
  sound_speed_body = sound_speed_sw * 1.0279,
  theta_body = pi / 2
)

bent_object <- fls_generate(
  shape = cylinder(
    length_body = 10.5e-3,
    radius_body = 1e-3,
    radius_curvature_ratio = 1.5,
    n_segments = 401
  ),
  density_body = density_sw * 1.0357,
  sound_speed_body = sound_speed_sw * 1.0279,
  theta_body = pi / 2
)
```

# Summary

```{r echo = FALSE}
bcms_summary <- utils::read.csv(
  file.path("..", "..", "scratch", "bcms_reference_compare_summary.csv")
)

knitr::kable(
  bcms_summary,
  digits = 6,
  col.names = c(
    "Case",
    "Max abs. delta TS (dB)",
    "Mean abs. delta TS (dB)",
    "f at max delta (kHz)",
    "acousticTS impl.",
    "acousticTS elapsed (s)",
    "f min (kHz)",
    "f max (kHz)",
    "Step (kHz)",
    "Reference impl.",
    "Reference elapsed (s)",
    "Ref. f min (kHz)",
    "Ref. f max (kHz)",
    "Ref. step (kHz)"
  )
)
```

Both branches agree exactly on the shared grid. That is the expected outcome for a model that is defined directly in terms of those two reference constructions.

## Spectrum overlays

```{r fig.width = 8, fig.height = 8}
bcms_compare <- utils::read.csv(
  file.path("..", "..", "scratch", "bcms_reference_compare.csv")
)
bcms_compare <- bcms_compare[order(bcms_compare$case, bcms_compare$frequency_hz), ]

straight_df <- subset(bcms_compare, case == "straight")
bent_df <- subset(bcms_compare, case == "bent")

par(mfrow = c(2, 2), mar = c(4, 4.5, 2, 1))

plot(
  straight_df$frequency_khz,
  straight_df$TS_reference,
  type = "l",
  lwd = 2.5,
  col = "black",
  xlab = "",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  main = "Straight cylinder"
)
lines(straight_df$frequency_khz, straight_df$TS_acousticts, col = "#0b6e4f", lwd = 2)

plot(
  straight_df$frequency_khz,
  straight_df$TS_acousticts - straight_df$TS_reference,
  type = "l",
  lwd = 2,
  col = "#0b6e4f",
  xlab = "Frequency (kHz)",
  ylab = expression(Delta * TS ~ (dB)),
  main = "Straight delta"
)
abline(h = 0, lty = 3, col = "grey50")

plot(
  bent_df$frequency_khz,
  bent_df$TS_reference,
  type = "l",
  lwd = 2.5,
  col = "black",
  xlab = "",
  ylab = expression(Target ~ strength ~ (dB ~ re. ~ 1 ~ m^2)),
  main = "Bent cylinder"
)
lines(bent_df$frequency_khz, bent_df$TS_acousticts, col = "#0b6e4f", lwd = 2)

plot(
  bent_df$frequency_khz,
  bent_df$TS_acousticts - bent_df$TS_reference,
  type = "l",
  lwd = 2,
  col = "#0b6e4f",
  xlab = "Frequency (kHz)",
  ylab = expression(Delta * TS ~ (dB)),
  main = "Bent delta"
)
abline(h = 0, lty = 3, col = "grey50")
```

# Closing note

This implementation page is intentionally narrow. `BCMS` is not trying to introduce a new scattering kernel. It is trying to reproduce a very specific relationship between a straight finite-cylinder modal sum and a bent-cylinder coherence correction. The exact agreement above shows that the code is honoring that structure.
