---
title: "Bent cylinder modal series solution"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Bent cylinder modal series solution}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  family = "bcms",
  pages = c(
    Overview = "index.html",
    Implementation = "bcms-implementation.html",
    Theory = "bcms-theory.html"
  )
)
```

The bent-cylinder modal series solution (`BCMS`) is a curvature-aware extension
of the straight finite-cylinder modal family developed by Stanton for
finite-length cylinders near broadside.[^1][^2] The physical idea is simple:
curvature modifies the way different parts of the cylinder remain coherent with
one another, but it does not replace the local cross-sectional scattering
physics with a completely different kernel.

That separation leads to a two-level theory:

1. a straight finite-cylinder modal backscatter kernel, and
2. a curvature-dependent coherent-length correction applied to that kernel.

The present page follows the package-wide indexing convention: medium `1` is
the surrounding seawater and medium `2` is the cylinder interior.

[^1]: Stanton, T.K. (**1988**). *Sound scattering by cylinders of finite length. I. Fluid cylinders*. *The Journal of the Acoustical Society of America*, 83: 55-63.

[^2]: Stanton, T.K. (**1989**). *Sound scattering by cylinders of finite length. III. Deformed cylinders*. *The Journal of the Acoustical Society of America*, 86: 691-705.

# Straight-cylinder starting point

## Local cross-sectional modal content

`BCMS` inherits its local cross-sectional response from the straight
finite-cylinder modal series. For a circular cylinder of radius $a$ and length
$L$, the near-broadside backscattering amplitude can be written schematically
as:

$$
  f_{\mathrm{bs}}^{(\mathrm{straight})} =
    \frac{L}{\pi}
  \frac{\sin(k_1 L \cos\theta)}
  {k_1 L \cos\theta}
  \sum_{m=0}^{\infty}
  (-1)^m \epsilon_m B_m.
$$

Here $k_1 = \omega/c_1$ is the seawater wavenumber, $\theta$ is the incidence
angle measured relative to the cylinder axis, $\epsilon_m$ is the Neumann
factor, and $B_m$ is the straight-cylinder modal coefficient of order $m$.

The details of $B_m$ depend on the boundary condition. For fluid-like cylinders,
those coefficients are the same ones derived in the
[FCMS theory page](../fcms/fcms-theory.html). `BCMS` does not replace those
coefficients. It reuses them.

## Why curvature can be isolated

For a gently and uniformly bent cylinder near broadside, the local radius and
cross-sectional boundary condition still look straight at the scale of the
cross-sectional modal solve. What changes is the two-way phase accumulated by
different points along the curved axis.

That is the central approximation of `BCMS`: curvature is treated as an
axial-coherence problem, not as a new cross-sectional boundary-value problem.

# Uniformly bent geometry

## Centerline and curvature

Let $s \in [-L/2, L/2]$ denote arc length along the cylinder centerline, and
let $\kappa = 1/\rho_c$ denote the constant curvature, where $\rho_c$ is the
radius of curvature. A convenient planar representation of the bent centerline
is:

$$
  \mathbf{r}_c(s) =
    \begin{bmatrix}
  \rho_c \sin(s/\rho_c) \\
    0 \\
    \rho_c \left[1 - \cos(s/\rho_c)\right]
  \end{bmatrix}.
$$

up to a rigid translation that does not affect the coherent integral.

The local tangent direction is then:

$$
  \hat{\mathbf{t}}(s) =
    \frac{d\mathbf{r}_c}{ds} =
    \begin{bmatrix}
  \cos(s/\rho_c) \\
    0 \\
    \sin(s/\rho_c)
  \end{bmatrix}.
$$

```{r echo = FALSE, out.width = "92%", fig.align = "center", fig.cap = "BCMS keeps the straight finite-cylinder modal sum and modifies only the along-axis coherence for a uniformly bent centerline."}
knitr::include_graphics("bcms-cylinder-schematic.svg")
```

## Broadside phase bookkeeping

For monostatic backscatter, each point on the centerline contributes a two-way
phase proportional to its projection onto the backscatter direction. If
$\hat{\mathbf{q}}$ denotes the relevant unit look direction, the coherent
length is:

$$
  L_{\mathrm{ebc}} =
    \int_{-L/2}^{L/2}
  \exp\left[
  2 i k_1 \hat{\mathbf{q}}\cdot \mathbf{r}_c(s)
  \right] ds.
$$

When the centerline is straight, $\mathbf{r}_c(s)$ becomes linear in $s$ and
this integral reduces to the ordinary sinc-style axial factor. For a bent
centerline, the phase becomes nonlinear in $s$, which is why the coherence is
reduced even when the local cylinder physics is unchanged.

# Equivalent coherent length and Fresnel form

For a uniformly bent cylinder near broadside, Stanton's reduction writes the
bent amplitude as:

$$
  f_{\mathrm{bs}}^{(\mathrm{bent})} =
    \frac{L_{\mathrm{ebc}}}{L}
  f_{\mathrm{bs}}^{(\mathrm{straight})}.
$$

The curvature therefore enters only through the ratio $L_{\mathrm{ebc}}/L$.
This quantity is often called the equivalent coherent length because it says
how much of the nominal straight-cylinder length remains phase coherent once
the axis is bent.

For constant curvature, the phase in the axial integral is quadratic in the
near-broadside reduction, so $L_{\mathrm{ebc}}$ can be expressed through
Fresnel integrals. The exact Fresnel form is useful computationally because it
avoids having to re-discretize the entire axis for each frequency. More
importantly, it makes the physics explicit: if curvature is weak or frequency
is low, the phase varies slowly and $L_{\mathrm{ebc}} \approx L$; if curvature
is stronger or frequency is higher, different portions of the bent axis
dephase and the coherent length decreases.

# Backscatter and target strength

Once the straight modal kernel and bent coherent-length factor are known, the
backscattering cross-section and target strength follow the standard monostatic
definitions:

$$
  \sigma_{\mathrm{bs}} =
    \left|f_{\mathrm{bs}}^{(\mathrm{bent})}\right|^2,
  \qquad
  \mathrm{TS} =
    10\log_{10}\left(\sigma_{\mathrm{bs}}\right).
$$

The curved cylinder therefore differs from the straight cylinder through a
complex coherence multiplier, not through a new definition of target strength.

# Mathematical assumptions

The family rests on a narrow but physically useful set of assumptions:

1. the cross-section remains circular,
2. the curvature is uniform,
3. the target is treated near broadside,
4. the straight-cylinder modal coefficients remain the correct local kernel,
5. curvature modifies only the axial phase coherence.

These assumptions are why `BCMS` is best understood as a curvature extension of
`FCMS`, not as a completely separate exact modal family. Its strength is that
it preserves the modal physics of the straight cylinder while still accounting
for the first-order way in which a bent axis destroys coherence.
