---
title: "Viscous-elastic spherical scattering model"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Viscous-elastic spherical scattering model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# acousticTS implementation

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  family = "vesm",
  pages = c(
    Overview = "index.html",
    Theory = "vesm-theory.html",
    Implementation = "vesm-implementation.html",
    Benchmarks = "vesm-benchmarks.html",
    Validation = "vesm-validation.html"
  )
)
```


These pages are motivated by layered gas-bearing fish scattering models and viscous resonance broadening [@Khodabandeloo_2021; @Feuillade_1998].

The viscous-elastic spherical model is available through `target_strength(..., model = "vesms")`. The implementation is built on top of `ESS` objects so that the gas core and elastic shell are stored in the same object framework already used by the elastic-shelled sphere model. The additional viscous layer is then supplied through model arguments at run time.

This page focuses on a single reproducible layered target based on the model presented by Khodabandeloo et al. (2021). The goal is to show how the gas core, elastic shell, and viscous layer are represented, how the model is called, and what result table is stored on the object.

::: {.experiment data-title="Evidence scope"}
`VESMS` has benchmark evidence and radial FEM validation for the documented layered-sphere case with matched geometry, material constants, modal orders, and frequency grid.
:::

## Building the reference object

The reference case uses:

- gas-core radius `R4 = 1.00 mm`
- shell outer radius `R3 = 1.02 mm`
- shell density `1040 kg m^-3`
- gas density `80 kg m^-3`
- gas sound speed `325 m s^-1`
- shell shear modulus `0.2 MPa`
- shell bulk modulus implied by `lambda = 2.4 GPa`
- viscous-layer density `1040 kg m^-3`
- viscous-layer sound speed `1510 m s^-1`
- viscous-layer shear and bulk viscosity both set to `3 kg m^-1 s^-1`
- surrounding seawater density `1027 kg m^-3`
- surrounding seawater sound speed `1500 m s^-1`

The shell and gas core are stored in an `ESS` object:

```{r}
library(acousticTS)

radius_gas <- 1e-3
radius_shell <- radius_gas + 0.02e-3
shear_shell <- 0.2e6
lambda_shell <- 2.4e9
bulk_shell <- lambda_shell + 2 * shear_shell / 3
sphere_shape <- sphere(radius_body = radius_shell, n_segments = 80)

vesm_object <- ess_generate(
  shape = sphere_shape,
  radius_shell = radius_shell,
  shell_thickness = radius_shell - radius_gas,
  density_shell = 1040,
  density_fluid = 80,
  sound_speed_fluid = 325,
  G = shear_shell,
  K = bulk_shell
)
```

## Running the model

The documented case retains the `m = 0, 1, 2` terms. In acousticTS, the corresponding setting is `m_limit = 2`. If no outer viscous radius is supplied, `VESMS` estimates it from the neutral-buoyancy relation described on the theory page.

```{r}
frequency <- seq(1e3, 150e3, by = 1e3)

vesm_object <- target_strength(
  object = vesm_object,
  frequency = frequency,
  model = "vesms",
  sound_speed_sw = 1500,
  density_sw = 1027,
  sound_speed_viscous = 1510,
  density_viscous = 1040,
  shear_viscosity_viscous = 3,
  bulk_viscosity_viscous = 3,
  m_limit = 2
)

head(extract(vesm_object, "model")$VESMS)
```

The stored output includes:

- `ka_viscous`, `ka_shell`, and `ka_gas`
- the complex backscattering amplitude `f_bs`
- the linear backscattering cross-section `sigma_bs`
- target strength `TS`

## Evidence reference

For VESM evidence, use:

- [VESM benchmarks](vesm-benchmarks.html)
- [VESM validation](vesm-validation.html)
