---
title: "Building Shapes"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Building Shapes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r echo=FALSE}
plot_shape_profile <- function(shape_obj, main) {
  pos <- extract(shape_obj, "position_matrix")
  x <- pos[, "x"]
  zU <- pos[, "zU"]
  zL <- pos[, "zL"]

  plot(x, zU,
    type = "l",
    ylim = range(c(zU, zL), finite = TRUE),
    xlab = "x",
    ylab = "z",
    main = main
  )
  lines(x, zL)
  abline(h = 0, lty = 3, col = "grey70")
}
```

# Overview

A shape describes geometry only. It does not assign material properties,
boundary conditions, or a physical target type. Those are added when the shape
is used to build a scatterer.

Every constructor on this page returns an S4 object derived from the
[`Shape` class](../../reference/Shape-class.html). A `Shape` contains a
`position_matrix` and a `shape_parameters` list. Keeping geometry separate
makes it possible to inspect, reuse, or simplify a body before choosing a
scattering model.

::: {.note data-title="Choose geometry for the intended model"}
A detailed outline is not automatically better. Use a canonical shape when its
geometry is part of the model assumption. Preserve a measured or segmented
shape when local morphology is important to the calculation.
:::

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_static_figure(
  src = "building-shapes-spectrum.png",
  alt = "Geometric representations ranging from canonical shapes to arbitrary segmented bodies"
)
```

# Choosing a constructor

| Constructor | Representation | Main inputs |
|---|---|---|
| [`sphere()`](../../reference/sphere.html) | Sphere | Radius |
| [`cylinder()`](../../reference/cylinder.html) | Straight or tapered cylinder | Length and radius, or a length-to-radius ratio |
| [`prolate_spheroid()`](../../reference/prolate_spheroid.html) | Elongated spheroid | Body length and maximum radius |
| [`oblate_spheroid()`](../../reference/oblate_spheroid.html) | Flattened spheroid | Body-axis length and equatorial radius |
| [`polynomial_cylinder()`](../../reference/polynomial_cylinder.html) | Cylinder with a polynomial radius profile | Length, radius, and polynomial coefficients |
| [`arbitrary()`](../../reference/arbitrary.html) | Measured or user-defined profile | Coordinate and dimension vectors |

The direct constructors make the resulting class clear. The
[`create_shape()`](../../reference/create_shape.html) wrapper can dispatch to a
canonical constructor when a single programmatic interface is more convenient.

# Canonical shapes

Canonical shapes are useful for model families derived in spherical,
cylindrical, or spheroidal coordinates. They are also useful for benchmarks and
controlled model comparisons. Construct them with dimensions in meters unless
the surrounding workflow explicitly handles another unit convention:

```{r}
library(acousticTS)

sphere_shape <- sphere(
  radius_body = 0.01,
  n_segments = 40
)

cylinder_shape <- cylinder(
  length_body = 0.05,
  radius_body = 0.003,
  n_segments = 80
)

prolate_shape <- prolate_spheroid(
  length_body = 0.04,
  radius_body = 0.004,
  n_segments = 60
)

oblate_shape <- oblate_spheroid(
  length_body = 0.012,
  radius_body = 0.01,
  n_segments = 60
)
```

`n_segments` controls the axial discretization. Increase it when the radius
profile or downstream numerical method requires finer resolution, then check
that the result is insensitive to further refinement.

```{r echo=FALSE, fig.width=8, fig.height=6, fig.cap="Profiles produced by the four canonical constructors."}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(2, 2), mar = c(3, 3, 2.2, 0.8))
plot_shape_profile(sphere_shape, main = "Sphere")
plot_shape_profile(cylinder_shape, main = "Cylinder")
plot_shape_profile(prolate_shape, main = "Prolate spheroid")
plot_shape_profile(oblate_shape, main = "Oblate spheroid")
par(old_par)
```

# Stored coordinates

Canonical and arbitrary shapes use a common axial profile. The `x` column gives
the axial location. The `zU` and `zL` columns give the upper and lower profile
bounds. Some arbitrary shapes also contain `w`, the transverse span in the
`y` direction.

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_static_figure(
  src = "building-shapes-coordinates.png",
  alt = "Stored coordinate conventions for spheres, cylinders, prolate spheroids, and arbitrary profiles"
)
```

For a sphere of radius $a$ centered at $x_c$, the stored profile follows:

$$
z_U(x) = \sqrt{a^2 - (x-x_c)^2},
\qquad
z_L(x) = -z_U(x).
$$

For a straight cylinder of radius $a$, the stored bounds are:

$$
z_U(x) = a,
\qquad
z_L(x) = -a.
$$

A polynomial or tapered cylinder retains the same columns while allowing the
local radius to vary with $x$.

For a prolate spheroid, the body semimajor axis lies along $x$ and the
semiminor axis lies along $z$. The corresponding focal half-distance is:

$$
q = \sqrt{a^2-b^2},
$$

Here, the semimajor axis $a$ is greater than the semiminor axis $b$. These
coordinates describe the geometry. Target orientation is assigned later when
building the scatterer.

Use [`extract()`](../../reference/extract.html) to inspect either slot without
editing the S4 object directly:

```{r}
head(extract(prolate_shape, "position_matrix"))
extract(prolate_shape, "shape_parameters")
```

# Arbitrary and measured shapes

Use `arbitrary()` when measured coordinates, taper, curvature, or anatomical
detail should be retained. It accepts an existing position matrix or named
coordinate vectors. The following symmetric profile is defined by axial
locations and local radii:

```{r}
measured_like <- arbitrary(
  x_body = c(0, 0.01, 0.02, 0.03, 0.04),
  radius_body = c(0, 0.004, 0.006, 0.004, 0)
)

extract(measured_like, "position_matrix")
```

For asymmetric profiles, provide `zU_body` and `zL_body`. A transverse
`w_body` vector can record width at the same axial stations. All coordinate
vectors should describe the same body in a consistent coordinate system and
unit convention.

```{r echo=FALSE, fig.width=7, fig.height=3.6, fig.cap="An arbitrary profile defined by axial locations and local radii."}
plot_shape_profile(measured_like, main = "Arbitrary measured-like body")
```

Operations such as translating, reanchoring, resampling, smoothing, and
inflating a shape are covered in [Shape
Manipulation](../shape-manipulation-reforging/shape-manipulation-reforging.html).

# Fitting a canonical surrogate

A measured body sometimes needs a canonical approximation for a shape-specific
model or a controlled comparison. This approximation should be explicit.
[`canonicalize_shape()`](../../reference/canonicalize_shape.html) can fit a
sphere, cylinder, prolate spheroid, or oblate spheroid while returning
diagnostics for the reduction.

Available fitting rules preserve volume, preserve length and volume, or fit the
equivalent-radius profile by least squares. The default depends on the requested
target shape. There is no single best rule for every measured body.

```{r}
cylinder_fit <- canonicalize_shape(
  measured_like,
  to = "Cylinder",
  diagnostics = TRUE
)

data.frame(
  fitted_shape = class(cylinder_fit$shape)[1],
  source_length_m = cylinder_fit$diagnostics$source$length,
  fitted_length_m = cylinder_fit$diagnostics$target$length,
  radius_rmse = cylinder_fit$diagnostics$fit$radius_rmse,
  radius_nrmse = cylinder_fit$diagnostics$fit$radius_nrmse
)
```

```{r echo=FALSE, fig.width=7, fig.height=3.6, fig.cap="The original arbitrary profile and its fitted cylindrical surrogate."}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(3, 3, 2.2, 0.8))
plot_shape_profile(measured_like, main = "Original arbitrary shape")
plot_shape_profile(cylinder_fit$shape, main = "Cylinder surrogate")
par(old_par)
```

Judge the surrogate using both the plotted profiles and the reported length,
volume, radius, and fit errors. A model applied to the fitted object describes
the canonical surrogate, not the original irregular body.

# Before building a scatterer

Check the following before adding physical properties:

- overall length and maximum radius
- axial ordering and coordinate units
- upper, lower, and transverse profiles where present
- segment resolution
- agreement between the geometry and the intended model assumptions
- canonicalization error if a surrogate was fitted

Next, [prepare the material properties](../material-properties/material-properties.html)
and use [Building Scatterers](../building-scatterers/building-scatterers.html) to
combine the geometry with its physical interpretation.
