---
title: "Working with Example Data"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Working with Example Data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

`acousticTS` includes three example scatterers and one collection of benchmark
results. They provide known inputs for learning package workflows, comparing
models, and checking calculations before working with custom targets.

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_static_figure(
  src = "example-data-map.png",
  alt = paste(
    "Bundled data support learning workflows, model comparison, and",
    "validation"
  )
)
```

# Included data

| Object | Contents | Primary use |
|---|---|---|
| [`sardine`](../../reference/sardine.html) | An `SBF` scatterer with separate body and swimbladder components, based on the NOAA KRM and `echoSMs` resources [@NOAA_KRM_software]. | Composite fish workflows. |
| [`cod`](../../reference/cod.html) | An `SBF` scatterer representing the Cod D case, with separate body and swimbladder components [@Clay_1994; @NOAA_KRM_software]. | Composite fish workflows and KRM comparisons. |
| [`krill`](../../reference/krill.html) | An `FLS` scatterer based on the krill geometry of McGehee et al. (1998) [@mcgehee_software]. | Fluid-like, elongated-body models. |
| [`benchmark_ts`](../../reference/benchmark_ts.html) | Stored target-strength spectra for canonical benchmark cases from Jech et al. (2015) [@Jech_2015; @echoSMs_software]. | Validation and regression checks. |

The first three objects are S4 scatterers that contain geometry, physical
properties, orientation, metadata, and containers for model parameters and
results. `benchmark_ts` is a list of reference outputs, not a scatterer.
The linked reference pages document each object's complete structure and
provenance.

# Loading and inspecting an object

Load a dataset explicitly with `data()`, then use the same inspection methods
available for other package objects:

```{r}
library(acousticTS)

data("krill", package = "acousticTS")
class(krill)
plot(krill)
```

Use `sardine` or `cod` to examine body-and-swimbladder targets. Use `krill` for
a single fluid-like body. Use `benchmark_ts` when reproducing a documented
benchmark under the same model assumptions, frequency grid, and numerical
settings.

::: {.note data-title="Example objects are starting points"}
These objects support examples and validation. They are not substitutes for
checking the geometry, material properties, orientation, and model assumptions
required by a new scientific application.
:::

# Related guides

- [Building Scatterers](../building-scatterers/building-scatterers.html)
- [Running Models](../running-models/running-models.html)
- [Validation and Benchmarks](../validation-benchmarks/validation-benchmarks.html)
