---
title: "Simulation and Parameter Sweeps"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Simulation and Parameter Sweeps}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

A single target-strength run answers a question about one specified target
state. [`simulate_ts()`](../../reference/simulate_ts.html) repeats that workflow
over deterministic parameter combinations, stochastic realizations, or both.
This is useful for orientation distributions, morphology studies, uncertain
material properties, and matched comparisons among models
[@Jech_2015; @Demer_2005].

![A baseline scatterer, parameter rules, the simulation design, and the collected model output.](simulation-batching-schematic.svg)

`simulate_ts()` calls `target_strength()` for each job and returns a named list
of data frames, one per model. Each row retains the realized parameters beside
the acoustic output.

# A minimal simulation

Begin with a valid `Scatterer`, then specify the frequency, model, and parameters
to vary:

```{r eval=FALSE}
library(acousticTS)

shape_obj <- cylinder(
  length_body = 0.05,
  radius_body = 0.003,
  n_segments = 80
)

obj <- fls_generate(
  shape = shape_obj,
  density_body = 1045,
  sound_speed_body = 1520
)

sim <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 6e3),
  model = "dwba",
  n_realizations = 5,
  parameters = list(
    theta_body = function() runif(1, 0.5 * pi, pi),
    density_body = 1045
  ),
  parallel = FALSE
)

head(sim$DWBA)
```

With no `parameters` and no `n_realizations`, the function performs one run of
the unchanged object. Use `target_strength()` directly when that is all the
workflow requires.

# How parameters are interpreted

The meaning of a parameter depends on how it is supplied:

![Fixed values, explicit vectors, and generating functions have different simulation roles.](simulation-parameter-modes.svg)

- A scalar is fixed across all jobs.
- A deterministic vector with more than one value defines a sweep axis.
- A function is called once for each realization and supplies a new value.
- A structured value, such as a named `body_target` vector, is passed to the
  applicable object-rebuilding pathway as one unit.

For example:

```{r eval=FALSE}
parameters <- list(
  sound_speed_body = 1520,
  theta_body = seq(0.5 * pi, pi, length.out = 5),
  density_body = function() rnorm(1, mean = 1045, sd = 5),
  body_target = c(length = 0.045)
)
```

This defines five deterministic orientation cells. Within each cell,
`density_body` is redrawn for every realization. Sound speed and target length
remain fixed.

::: {.important data-title="Vectors are sweep axes"}
A multi-value vector is not matched automatically to `n_realizations`. It
defines a deterministic axis. Use a generating function for realization-level
randomness or `permute = FALSE` to pair several deterministic axes.
:::

# Design cells and realizations

Every deterministic multi-value parameter becomes an outer design axis.
`n_realizations` controls the number of repeated evaluations inside each design
cell:

![Cartesian parameter cells outside repeated within-cell realizations.](simulation-batch-cells.svg)

The following design has four orientations, three densities, and three
realizations per cell. It therefore creates 36 jobs for each requested model:

```{r eval=FALSE}
sim_grid <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 6e3),
  model = c("dwba", "sdwba"),
  n_realizations = 3,
  parameters = list(
    theta_body = seq(0.5 * pi, pi, length.out = 4),
    density_body = c(1035, 1045, 1055)
  ),
  batch_by = c("theta_body", "density_body"),
  parallel = FALSE
)
```

Multi-value parameters are detected automatically, so `batch_by` is not
required to make those vectors vary. Naming them in `batch_by` makes the outer
design explicit and is useful when reading or generating a simulation call.

## Crossing or pairing axes

With `permute = TRUE`, the default, deterministic axes form a Cartesian grid.
Set `permute = FALSE` when their values describe matched states that should
advance together:

```{r eval=FALSE}
sim_paired <- simulate_ts(
  object = obj,
  frequency = 70e3,
  model = "dwba",
  parameters = list(
    theta_body = seq(0.5 * pi, pi, length.out = 4),
    density_body = c(1035, 1045, 1050, 1055)
  ),
  permute = FALSE,
  parallel = FALSE
)
```

This creates four paired states rather than sixteen crossed states. All varied
axes must have the same length. Length-one inputs remain fixed.

# Parameters that rebuild the target

Some parameters alter the target rather than only the model call. Compatible
arguments are applied through [`reforge()`](../../reference/reforge.html), and
matching component fields can be updated before the model is run. Each output
then represents a distinct target state derived from the baseline object:

![A baseline scatterer is reforged into distinct geometric states before each model run.](simulation-object-reforge.svg)

For an `FLS` object, convenience names such as `length_body` use the same
rebuilding pathway as `body_target`:

```{r eval=FALSE}
sim_length <- simulate_ts(
  object = obj,
  frequency = 120e3,
  model = "dwba",
  n_realizations = 4,
  parameters = list(
    length_body = c(0.04, 0.05, 0.06),
    theta_body = function() runif(1, 0.5 * pi, pi)
  ),
  parallel = FALSE
)
```

Use a list when each sweep level is itself a structured target specification:

```{r eval=FALSE}
sim_shape <- simulate_ts(
  object = obj,
  frequency = 120e3,
  model = "dwba",
  parameters = list(
    body_target = list(
      c(length = 0.04, radius = 0.0025),
      c(length = 0.06, radius = 0.0035)
    ),
    isometric_body = FALSE
  ),
  parallel = FALSE
)
```

For an already bent `FLS` object, `length_body` refers to the bent centerline arc
length. `simulate_ts()` does not bend a straight baseline automatically. Create
the intended baseline with [`brake()`](../../reference/brake.html) before the
simulation. If bending must vary by realization, construct those objects
outside `simulate_ts()` so the order of resizing and bending is explicit.

# Inspect and summarize output

The returned object is organized by model:

```{r eval=FALSE}
names(sim_grid)
head(sim_grid$DWBA)
head(sim_grid$SDWBA)
```

Group summaries should respect the design columns and `realization` index.
Common summaries include mean or quantile target strength by frequency,
orientation distributions, sensitivity across morphology cells, and matched
differences between models. Perform energetic averages in a linear
cross-section domain rather than averaging target strength in decibels unless a
specific reporting convention requires otherwise.

# Parallel execution and reproducibility

The parallel pathway distributes the same logical job table over PSOCK workers:

![Serial and parallel execution use the same simulation jobs with different scheduling.](simulation-execution-layout.svg)

Use `parallel = FALSE` while developing a design. PSOCK startup and, when
working from a source checkout, preparation of a worker installation can cost
more than the model calls for a small job. Parallel execution becomes more
useful once individual jobs are sufficiently expensive.

```{r eval=FALSE}
sim_parallel <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 2e3),
  model = "dwba",
  n_realizations = 100,
  parameters = list(
    theta_body = function() runif(1, 0.5 * pi, pi)
  ),
  parallel = TRUE,
  n_cores = 4
)
```

Record the parameter definitions, package version, and random-number settings
for reproducible stochastic work. Verify a stochastic design serially before
assuming that worker scheduling will reproduce the same draws as a sequential
run.

# Plan the grid before running it

For crossed axes, the number of jobs per model is the product of the axis
lengths multiplied by `n_realizations`. Frequency points add rows to each job's
result. Several moderate axes can therefore create a large output and a large
number of model calls.

Before a production run:

1. calculate the number of design cells and expected output rows,
2. run a small serial subset and inspect the realized parameters,
3. confirm that crossed or paired semantics match the scientific design,
4. verify that geometry-changing parameters rebuild the intended target, and
5. split very large studies into recoverable batches or a cached project-level
   pipeline.

`simulate_ts()` is appropriate for a structured family of related model calls.
For a long-running study with external data dependencies, many intermediate
products, or expensive downstream summaries, a workflow manager such as
`targets` can orchestrate separate `simulate_ts()` calls without becoming a
dependency of `acousticTS`.
