---
title: "Simulation and Parameter Sweeps"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Simulation and Parameter Sweeps}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

Single deterministic model runs are useful, but many acoustic questions are inherently distributional. The package therefore includes simulation tools for repeated realizations, parameter batching, and ensemble summaries.

![Batching and realizations schematic](simulation-batching-schematic.svg)

This page is about how to turn a single target description into a structured family of model runs. The core idea is that the object remains fixed as the bookkeeping container, while selected inputs are varied across realizations or across a parameter grid.

# Main entry point

The main current simulation interface is `simulate_ts()`. It supports repeated draws of user-specified parameters, optional batching over one or more variables, and optional parallel execution.

In other words, `simulate_ts()` is a workflow wrapper around repeated `target_strength()` calls. It does not replace the usual model layer. It automates the process of rebuilding a set of parameters, optionally reworking the object, running the model, and collecting the outputs into a set of tidy result tables.

```{r eval = TRUE}
library(acousticTS)

shape_obj <- cylinder(
  length_body = 0.05,
  radius_body = 0.003,
  n_segments = 80
)

obj <- fls_generate(
  shape = shape_obj,
  density_body = 1045,
  sound_speed_body = 1520
)

res <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 6e3),
  model = "DWBA",
  n_realizations = 5,
  parameters = list(
    theta_body = function() runif(1, min = 0.5 * pi, max = pi),
    density_body = 1045
  ),
  parallel = FALSE
)
```

The return value is a named list with one data frame per model.

# Parameter specification

The scratch simulation notes are worth keeping in mind here because they clarify the three main parameter modes supported by `simulate_ts()`:

1. single fixed values,
2. explicit vectors,
3. generating functions that return one draw at a time.

When `batch_by` is supplied, the selected parameters define a Cartesian grid and each grid cell is then repeated for `n_realizations`. That distinction between batch cells and within-cell draws is the core conceptual point of the interface.

Those three modes answer different questions:

1. fixed values are for constants that should be repeated in every realization,
2. explicit vectors are for a known set of realization-level values,
3. functions are for stochastic draws generated on demand.

The distinction matters because it separates deterministic sweeps from uncertainty propagation.

![Three parameter-specification modes in `simulate_ts()`: fixed constants, explicit vectors, and per-realization generating functions.](simulation-parameter-modes.svg)

Fixed values are copied into every run unchanged. Explicit vectors supply a known sequence of values. Generating functions do not provide a pre-listed table at all. They create a new draw whenever a realization is built. That difference is what separates a design grid from on-the-fly uncertainty propagation.

```{r eval = TRUE}
parameters <- list(
  theta_body = seq(0.5 * pi, pi, length.out = 5),
  density_body = function() rnorm(1, mean = 1045, sd = 5),
  sound_speed_body = 1520
)
```

In this example, `theta_body` is an explicit set of values, `density_body` is random, and `sound_speed_body` is held fixed.

# Batching versus realizations

The most important conceptual distinction in the interface is the difference between a batched parameter and a realization-level parameter.

1. `batch_by` creates a grid of settings that should each be kept distinct in the output.
2. `n_realizations` controls how many repeated runs are performed within each grid cell.

Suppose `theta_body` and `density_body` are both placed in `batch_by`. Then the simulation does not treat them as uncertain draws around one target. It treats them as a design grid whose combinations are all run separately.

```{r eval = TRUE}
res_grid <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 6e3),
  model = c("DWBA", "SDWBA"),
  n_realizations = 3,
  parameters = list(
    theta_body = seq(0.5 * pi, pi, length.out = 4),
    density_body = c(1035, 1045, 1055)
  ),
  batch_by = c("theta_body", "density_body"),
  parallel = FALSE
)
```

Here the output contains one row structure per model, per frequency, per batch cell, and per realization. That is why batched jobs can grow quickly.

![How `batch_by` creates Cartesian design cells and `n_realizations` nests repeated runs inside each cell.](simulation-batch-cells.svg)

This is the core simulation structure in the current interface. Batched parameters define the outer design. Realizations live inside that design. If a user loses track of that nesting, it becomes very easy to misread deterministic design cells as random draws or random draws as deterministic sweep levels.

# Parameters that modify the object

Not every simulated parameter is a simple scalar passed into `target_strength()`. Some parameters alter the working object itself. This is especially important for reshaping workflows.

The current simulation code supports two broad cases:

1. parameters matching accepted `reforge()` inputs can trigger per-realization resizing or reparameterization,
2. parameters matching object-component fields can be inserted into the working object before the model call.

This means `simulate_ts()` can be used not just for orientation or density perturbations, but also for controlled morphology studies.

```{r eval = TRUE}
body_shape <- arbitrary(
  x_body = c(0.00, 0.02, 0.05, 0.08),
  zU_body = c(0.001, 0.004, 0.004, 0.001),
  zL_body = c(-0.001, -0.004, -0.004, -0.001)
)

bladder_shape <- arbitrary(
  x_bladder = c(0.02, 0.04, 0.06),
  zU_bladder = c(0.0012, 0.0018, 0.0012),
  zL_bladder = c(-0.0012, -0.0018, -0.0012)
)

obj_sbf <- sbf_generate(
  body_shape = body_shape,
  bladder_shape = bladder_shape,
  density_body = 1045,
  sound_speed_body = 1520,
  density_bladder = 1.2,
  sound_speed_bladder = 343
)

res_shape <- simulate_ts(
  object = obj_sbf,
  frequency = seq(38e3, 120e3, by = 6e3),
  model = "DWBA",
  n_realizations = 2,
  parameters = list(
    body_scale = c(0.9, 1.1),
    swimbladder_inflation_factor = c(0.8, 1.2)
  ),
  batch_by = c("body_scale", "swimbladder_inflation_factor"),
  parallel = FALSE
)
```

This is one of the cleanest ways to keep a morphology-sensitivity analysis tied to the same baseline target.

![Object-modifying simulation workflow, where baseline geometry is reforged into a family of derived target states before model execution.](simulation-object-reforge.svg)

Morphology-affecting parameters do not behave like ordinary scalar metadata. They change the working target itself before the model is run. The resulting simulations should therefore be interpreted as distinct geometric states derived from one baseline object, not as repeated evaluations of a single unchanged shape.

# When simulation is useful

Simulation or batched evaluation is most useful when:

1. orientation is uncertain,
2. material properties are better treated as distributions than as fixed values,
3. multiple body lengths or frequencies must be screened systematically,
4. model sensitivity is of greater interest than a single nominal prediction.

In practice, this means simulation is less about "making the model stochastic" and more about mapping the consequences of uncertainty, variability, or design decisions.

# Interpreting the output

The returned list is model-centered. Each element is a data frame that carries the realization index, any batched parameters, and the model outputs. That organization is deliberate: it makes downstream plotting or aggregation straightforward.

```{r eval = TRUE}
names(res_grid)
head(res_grid$DWBA)
head(res_grid$SDWBA)
```

Common downstream summaries include:

1. mean and quantiles of `TS` by frequency,
2. variance decomposition by orientation or material parameter,
3. pairwise model differences across matched grid cells,
4. screening plots to identify nonlinear regions before deeper analysis.

# Parallel execution on Windows

On Windows, parallel execution uses PSOCK workers rather than forked processes. That has two practical consequences:

1. startup overhead is more noticeable for small jobs,
2. larger simulations benefit more clearly from parallelization than tiny exploratory runs.

For this reason, `parallel = FALSE` is often the right default while debugging a workflow, even if `parallel = TRUE` is the right choice for production runs.

```{r eval = FALSE}
res_parallel <- simulate_ts(
  object = obj,
  frequency = seq(38e3, 120e3, by = 2e3),
  model = "DWBA",
  n_realizations = 100,
  parameters = list(theta_body = function() runif(1, 0.5 * pi, pi)),
  parallel = TRUE,
  n_cores = 4
)
```

![Execution layout for serial and parallel simulation runs, emphasizing the same logical job table with different scheduling.](simulation-execution-layout.svg)

Parallelization changes scheduling, not interpretation. The same batched cells and realization-level jobs still exist. Parallel execution simply distributes them across workers. That is why debugging is often easier in serial mode first, even when the final production run is parallel.

# Practical cautions

The simulation workflow can become expensive quickly, especially when several parameters are batched simultaneously. The main practical risks are:

- combinatorial growth in the evaluation grid,
- unnecessary parallel overhead for small jobs,
- confusion between realization-level randomness and batch-level parameter sweeps.

Two additional cautions are worth keeping in mind:

1. if a parameter sequence is meant to represent uncertainty, placing it in `batch_by` changes the interpretation from random variation to deterministic design cells,
2. if geometry-affecting parameters are varied, the results should be interpreted as new target states rather than as mere perturbations of one fixed shape.

For that reason, this page should be read together with [FAQ and troubleshooting](../faq-troubleshooting/faq-troubleshooting.html).

On Windows, parallel execution uses PSOCK workers rather than forked processes, so larger simulations benefit from being planned deliberately instead of relying on a maximal batch grid by default.

# Relationship to `anneal()`

The current documentation and source code indicate that `simulate_ts()` is expected to give way to `anneal()` in future releases. The important point for current users is that the workflow logic on this page remains useful even if the interface evolves.

The stable ideas are:

1. keep one object as the baseline target description,
2. separate batch design from within-cell variation,
3. make model comparison operate on matched simulation cells,
4. summarize in linear or logarithmic units according to the scientific question.

# Relationship to future interfaces

The current simulation interface is already useful, but the package documentation indicates that this part of the workflow may continue to evolve. This article should therefore be read as a workflow guide rather than as a claim that the current interface is final.
