---
title: "Preparing Material Properties"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    dpi: 200
    dev.args: list(pointsize=11)
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Preparing Material Properties}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

Use this guide after choosing a target geometry and before building a
scatterer. It covers how to record, convert, and check the material values that
will be attached to the target. Constructor selection and object assembly are
covered in [Building Scatterers](../building-scatterers/building-scatterers.html).

::: {.note data-title="Physical interpretation"}
See [Scattering Boundary Conditions](../boundary_conditions.html) for how
material properties enter the interface conditions.
:::

```{r echo=FALSE, results='asis'}
acousticTS:::.vignette_static_figure(
  src = "material-property-pipeline.svg",
  alt = "Material property pipeline"
)
```

# Record the source values

Record each value with its units, source, and measurement conditions. Density
and sound speed can depend on temperature, salinity, pressure, and composition.
In dispersive or lossy media, effective wave properties may also depend on
frequency. A contrast copied from a publication is usable only when its
reference medium is known [@Medwin_1998].

::: {.note data-title="Symbols and units"}
See [Notation and Symbols](../notation-and-symbols/notation-and-symbols.html)
for the shared material-property definitions and units.
:::

Convert units before constructing the scatterer. In particular, elastic
moduli should be supplied in pascals and geometric dimensions should follow
the units declared in the constructor call.

# Make the reference medium explicit

Absolute target properties remain properties of the target, but their acoustic
interpretation and all derived ratios depend on the exterior medium used in the
calculation. Keep those medium values with the target setup, even when they are
not arguments to the scatterer constructor:

```{r}
library(acousticTS)

medium <- data.frame(
  density = 1026,
  sound_speed = 1500
)

body <- data.frame(
  density = 1045,
  sound_speed = 1520
)
```

The same exterior density and sound speed should later be supplied to the
scattering calculation. Otherwise, ratios derived during target preparation
and properties completed during model evaluation can describe different
physical systems.

# Choose absolute properties or ratios

Fluid-like constructors commonly accept either absolute values such as
`density_body` and `sound_speed_body` or directional ratios such as `g_body`
and `h_body`. For a body in the exterior medium, those ratios are
$g_{21}=\rho_2/\rho_1$ and $h_{21}=c_2/c_1$.

Compute the ratios from the recorded absolute values when both are available:

```{r}
g_body <- body$density / medium$density
h_body <- body$sound_speed / medium$sound_speed

data.frame(g_body, h_body)
```

Supply one representation for each property. Providing an absolute density
and a contradictory density ratio does not describe a unique target. The same
rule applies to sound speed.

Ratios and perturbation contrasts are also different quantities. The package
helpers `rho()` and `compressibility()` calculate the density and
compressibility perturbations used by weak-scattering formulations:

```{r}
data.frame(
  gamma_rho = acousticTS::rho(medium, body),
  gamma_kappa = compressibility(medium, body)
)
```

The ratios $g_{21}$ and $h_{21}$ approach one for a weak contrast. The
perturbations $\gamma_\rho$ and $\gamma_\kappa$ approach zero. They should not
be substituted for one another.

# Check combined fluid properties

Density and sound speed should also be checked together. Acoustic impedance
and absolute compressibility provide compact diagnostics for an interface:

```{r}
property_check <- rbind(
  exterior = medium,
  body = body
)

property_check$impedance <-
  property_check$density * property_check$sound_speed
property_check$compressibility_pa_inverse <-
  1 / (property_check$density * property_check$sound_speed^2)

property_check
```

A density ratio close to one does not guarantee a small acoustic contrast if
the sound speeds differ materially. Likewise, a gas component should not be
treated as weakly contrasting merely because the chosen density value has not
yet been checked against its sound speed.

# Convert elastic properties

Elastic properties are reported in several equivalent parameterizations. The
package provides `pois()`, `bulk()`, `young()`, `shear()`, and `lame()` to
convert between the common isotropic elastic constants [@Achenbach_1973].

For example, Young's modulus and Poisson ratio can be converted before an
elastic scatterer is built:

```{r}
E <- 70e9
nu <- 0.32
density_solid <- 2700

G <- shear(E = E, nu = nu)
K <- bulk(E = E, nu = nu)
lambda <- lame(E = E, nu = nu)

c_L <- sqrt((lambda + 2 * G) / density_solid)
c_T <- sqrt(G / density_solid)

data.frame(E, nu, G, K, lambda, c_L, c_T)
```

Supply only the independent values intended for the conversion. Redundant and
inconsistent elastic constants can hide which pair was actually used. For a
stable isotropic linear-elastic material, check that $K>0$, $G>0$, and $E>0$
as applicable, and that $-1<\nu<0.5$ before construction. Lamé's first
parameter need not be positive for every stable auxetic material.

Different target classes expose different elastic inputs. An elastic-shell
constructor may accept a pair such as `E` and `nu`, while a calibration sphere
may be specified by density and longitudinal and transverse wave speeds. Use
the representation supported by the intended constructor rather than
converting values that are already available in the required form.

# Keep layered components separate

For a layered target, retain one labeled record per physical region. Do not
collapse shell, core, viscous layer, and exterior values into an unlabeled set
of generic properties.

| Region | Values to retain |
|---|---|
| exterior fluid | density and sound speed |
| fluid-like body or core | density and sound speed, or documented ratios |
| elastic solid or shell | density and an independent elastic parameter set |
| viscous layer | density, compressional sound speed, shear viscosity, and bulk viscosity when required |

The theoretical ratios between adjacent media can always be derived from
these labeled absolute values. Constructor argument names may instead express
a component relative to the exterior medium, so check the constructor
documentation before supplying a precomputed ratio.

# Build a constructor-ready target

Once the values are checked, attach them to a compatible shape. Absolute
properties can be supplied directly:

```{r}
body_shape <- prolate_spheroid(
  length_body = 0.04,
  radius_body = 0.004,
  n_segments = 60
)

body_target <- fls_generate(
  shape = body_shape,
  density_body = body$density,
  sound_speed_body = body$sound_speed
)

extract(body_target, "body")
```

The same shape can instead be constructed from ratios derived against the
recorded exterior medium:

```{r}
body_target_ratios <- fls_generate(
  shape = body_shape,
  g_body = g_body,
  h_body = h_body
)

extract(body_target_ratios, "body")
```

These objects describe the same material only when the later model run uses
the exterior values stored in `medium`.

# Validation checklist

Before moving to scatterer construction or model evaluation, confirm that:

1. every dimensional value has been converted to the units expected by the package,
2. the exterior medium used to derive ratios is recorded,
3. absolute values and ratios do not contradict one another,
4. $g$ and $h$ ratios have not been confused with perturbation contrasts,
5. each region of a layered target has its own labeled property set,
6. elastic constants form a physically admissible independent set, and
7. frequency-dependent or complex properties have not been replaced silently by unrelated static values.

# Next step

Continue to [Building Scatterers](../building-scatterers/building-scatterers.html)
to select the target class and attach these values to the geometry.

# Related pages

- [Building Shapes](../building-shapes/building-shapes.html)
- [Building Scatterers](../building-scatterers/building-scatterers.html)
- [Scattering Boundary Conditions](../boundary_conditions.html)
- [SOEMS calibration spheres](../calibration/index.html)
