---
title: "BBFM"
output:
  rmarkdown::html_vignette:
    toc: false
    number_sections: false
    fig_caption: yes
bibliography: ../REFERENCES.bib
link-citations: true
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{BBFM}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Overview

```{r model_family_header, echo=FALSE, results='asis'}
acousticTS:::.model_family_header(
  family = "bbfm",
  pages = c(
    Overview = "index.html",
    Implementation = "bbfm-implementation.html",
    Theory = "bbfm-theory.html"
  )
)
```


This family is best read alongside the swimbladder-less fish and composite-scatterer literature that motivates explicit flesh-body and backbone terms [@gorska_etal_2005; @stanton_sound_1998-1; @clay_horne_1994].

```{r bbfm_overview, echo=FALSE, results='asis'}
acousticTS:::.model_family_overview(
  family = "bbfm",
  summary = paste(
    "The body-backbone fish model (`BBFM`) is the package's composite",
    "swimbladder-less family for targets whose flesh body and backbone should",
    "remain explicit acoustic components."
  ),
  core_idea = paste(
    "Compute a weak-fluid flesh-body term with `DWBA`, compute an elastic",
    "backbone term with `ECMS`, place the backbone inside the same body-fixed",
    "frame with a two-way phase factor, and then sum the two complex",
    "amplitudes coherently."
  ),
  best_for = c(
    "Swimbladder-less fish where the backbone should remain acoustically explicit",
    "Composite body-plus-backbone studies that are too structured for a body-only approximation",
    "Intermediate modeling between weak-fluid body models and future fully coupled composite solvers"
  ),
  supports = c(
    "`BBF` scatterers built from explicit `body_shape` and `backbone_shape` inputs",
    "Flesh-body material properties as contrasts or absolute density/sound speed",
    "Backbone density plus longitudinal and transverse elastic wave speeds"
  ),
  assumptions = c(
    "Flesh body treated in the weak-fluid `DWBA` regime",
    "Backbone treated as an elastic cylinder through `ECMS`",
    "Components combined coherently after centroid-based phase placement",
    "No repeated rescattering or fully coupled embedded elastic-cylinder solve"
  ),
  reading = c(
    "[Implementation](bbfm-implementation.html): object setup, stored outputs, and internal reconstruction checks",
    "[Theory](bbfm-theory.html): composite amplitude, phase placement, and interference structure"
  )
)
```

::: {.experiment data-title="Current family status"}
`BBFM` is already useful as a transparent composite family, but it should still
be treated as experimental. The main open work is external validation and, in
the longer term, a more tightly coupled treatment of the backbone as an
embedded elastic structure rather than a positioned component surrogate.
:::

## Why it exists

`BBFM` fills the gap between:

- body-only weak-scattering models that ignore the backbone,
- and future fully coupled body-plus-backbone solvers that are not yet in the
  package.

That is why the family belongs in acousticTS even before a more complete
composite solver exists: it gives a physically interpretable way to keep the
two dominant anatomical components explicit in one coherent target-strength
calculation.
